---
title: "Abundance modelling - LOB depth"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```



# Packages
```{r message=FALSE, warning=FALSE, include=FALSE}
source("code/packages.R")
#source('code/helper_functions.R')
source('code/extras.R')
```

# Load model and data
```{r}
load(file = "data/Rdata/mod_data/abun_depth_mod_final.Rdata")
load(file = "data/Rdata/lob_predsum.Rdata")
```

# Model summary
```{r}
abun_depth_mod_final
```
We need an accurate number for scaleDepth_m for plotting later (the probability density)

# Can do this as follows
```{r}
#get_variables(abun_depth_mod_final) # If you need to figure out exactly how the variable listed in the model

# Then
abun_depth_mod_final %>% 
  spread_draws(b_scaleDepth_m) %>% 
  summarise(median_hdci(b_scaleDepth_m))
```
-0.42

# Model investigation 
## Conditional effects plot again to aid interpretation
```{r}
abun_depth_mod_final |> conditional_effects() |> plot(ask = FALSE, points = FALSE) # Without raw data
#abun_depth_mod_final |> conditional_effects() |> plot(ask = FALSE, points = TRUE) # With raw data overplotted
```

## Summary table
```{r}
abun_depth_mod_final |> 
  as_draws_df() |>    # Extract draws
  exp() |>            # Back transform
  summarise_draws(median,
                  HDInterval::hdi,
                  rhat, ess_bulk, ess_tail,
                  Pl = ~mean(.x <1),                       # Exceedence prob - on exp scale it's sig if doesn't cross 1
                  Pg = ~mean(.x >1))|> 
  as_tibble() |> 
  mutate(median = round(median, 3),
         lower = round(lower, 3),
         upper = round(upper, 3))
```

### Interpretation:

- Strong evidence for a negative effect of depth:
0.658 * as many predators per 1SD increase in depth (not in actual meters units) 
-> Median = 1-0.658 equivalent to a 34.2% decline per 1SD
-> Max = 1-0.578 = 42.2%
-> Min = 1-0.747 = 25.3%


## emmeans and contrasts
Will compare abundance at max and min values of both predictors

```{r}
lob_predsum |> 
  summarise(min_Depth = min(Depth_m),
            max_Depth = max(Depth_m))
```

# Make dataframe with data points:
```{r}
depth_newdata <- data.frame(Depth_m = c(4, 99))
```


## emmeans for each extreme
```{r}
abun_depth_mod_final |> 
  emmeans(~ Depth_m, at = depth_newdata, type = "response") |> 
  as_tibble()
```
## So on average we have about 4 predators per transect in the shallows and 1 per transect at depth. 4:1 ratio

# Get contrast ratio ±HDI
```{r}
abun_depth_mod_final |> 
  emmeans(~ Depth_m, at = depth_newdata) %>% 
  pairs(reverse = FALSE) |> 
  gather_emmeans_draws() |> 
  summarise(median_hdci(exp(.value), .width = c(0.8, 0.95)), 
            Pl = mean(.value < 0),
            Pg = mean(.value > 0))

```
So over 4x as many predators in the shallows compared to the deepest depths surveyed.


## What is the rate of decline? Say per 10m depth increase..

Well, we can do a contrast between the min and max 
```{r}
abun_depth_mod_final |> 
  emmeans(~ Depth_m, at = depth_newdata, type = "link")  |> 
  regrid() |> 
  pairs(reverse = FALSE) |> 
  as_tibble()
```
99 - 4 = 95
95/3.17 = 30

## -> So a decline of about one predator per transect for every 30 meters of depth. 


# Alternative way of doing this:

```{r}
depth_newdata2 <- data.frame(Depth_m = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100))
```


```{r}
abun_depth_mod_final |> 
  emmeans(~ Depth_m, at = depth_newdata2, type = "response") %>% 
  as_tibble() |> 
  mutate(diff = prob - lag(prob))  |> 
  slice_tail(n = 10) |> 
  summarise(median_hdci(diff))

```


Absolute numbers not that useful here (as they relate to individual transect). Will just go with the contrast ratio:

4.5 more predatory fishes at 99 than 4 m (± HDCI 2.7 - 6.9)

# Interpreting summary table some more..
0.658 * as many predators per 1SD increase in depth (not in actual meters units)
So what is 1SD of depth?

```{r}
lob_predsum |> 
  summarise(mean_Depth = mean(Depth_m),
            sd_Depth = sd(Depth_m))
```

So a decline of 34.2% per 26.37m. Does this sound right?
Check difference between 10-36.4

```{r}
depth_newdata3 <- data.frame(Depth_m = c(40, 66.37))
```


```{r}
abun_depth_mod_final |> 
  emmeans(~ Depth_m, at = depth_newdata3, type = "link")  |> 
  regrid() |> 
  pairs(reverse = FALSE) |> 
  as_tibble()
```

A drop of 0.786 from 2.327 -> 1.54

2.327 * 0.658 = 1.53



## Figures

## Effect of depth
```{r}
depth_grid <- with(lob_predsum, list(Depth_m = seq(from = min(Depth_m), to = max(Depth_m), len = 100)))
```

# Extract posterior draws to reconstruct trend
```{r fig.height=70/26.5, fig.width=100/25.4}
depth_abun_trend <- 
abun_depth_mod_final |> 
  emmeans(~ Depth_m, epred = TRUE,
           at = depth_grid) |> 
  gather_emmeans_draws() |> 
  ggplot(aes(x = Depth_m, y = .value)) +
  #geom_point(data = lob_predsum %>% filter(No_Preds <12), aes(x = Depth_m, y = No_Preds), fill = "#0066cc", colour = "midnightblue", pch = 21, alpha = 0.3, position = position_jitter())+
  geom_line(aes(group = .draw), colour = "#0066cc", alpha = 0.05) +
  stat_lineribbon(colour = "midnightblue", size = 0.5, point_interval = "median_hdci", fill = "#0066cc", .width = 0.001, alpha = 1) +
  scale_y_continuous(name = expression("No. preds (.150m"^{2}~")"), limits = c(0,10), expand = c(0.05,0), breaks = c(0,2,4,6,8,10)) +
  scale_x_continuous("Depth (m below sea surface)") + 
  theme_bjc() +
  theme(axis.title.x = element_text(colour = "black", size = 10),
        axis.title.y = element_text(colour = "black", size = 10),
        axis.text = element_text(colour = "black", size = 9),
        legend.position = "none")

depth_abun_trend
```





# Probability distribution of effect size
```{r fig.width = 70/25.4, fig.height=50/25.4}
depth_abun_effect <- 
abun_depth_mod_final |> 
  gather_draws(`^b_.*`, regex=TRUE)  |>    # Need to use gather_draws here if you want to use the regex
  filter(.variable == 'b_scaleDepth_m') |>
  ungroup() |> 
  ggplot(aes(x = 1-exp(.value))) + 
  stat_halfeye(fill = "#0066cc", .width = c(0.8, 0.95), point_interval = "median_hdi") +
  geom_vline(xintercept = 0, lty = 11) +
  #geom_vline(xintercept = 0.342, colour = "white") + # 0.342 = median decline (%) per 1 SD depth (see summary table)
  #annotate(geom = "text", label = "P=1", x = -0.6, y = 0.8, colour = "black", size = 3.5, family = "Arial", hjust = 0) +
  scale_y_discrete(expand = c(0.08,0)) +
  scale_x_continuous("Scaled effect distribution") +
  theme_bjc() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(colour = "black", size = 10),
        axis.text.y = element_blank(),
        axis.text.x = element_text(colour = "black", size = 9),
        legend.position = "none")

depth_abun_effect
```

Median effect size is 34.2% per 1SD increase in depth 
25.3 - 42.2%


# Combine
```{r fig.width = 152/25.4, fig.height=60/25.4}
depth_abun_fig <- 
  cowplot::plot_grid(depth_abun_trend, depth_abun_effect, 
                     nrow = 1, labels = c("a", "b"), rel_widths = c(1.2, 0.6), label_size = 10, align = "h") +
  theme(plot.background = element_rect(color = "black"))
depth_abun_fig
```


# Save
```{r}
ggsave(filename = "output/depth_abun_fig.pdf", plot = depth_abun_fig, width = 152, height = 60, device = cairo_pdf, units = "mm", dpi = 300)
ggsave(filename = "output/depth_abun_fig.jpg", plot = depth_abun_fig, width = 152, height = 60, device = "jpeg", units = "mm", dpi = 300)
```




# Contrast figure: shallow to deep

# Compare across actual range of depths surveyed 4 - 99m
```{r}
depth_newdata2 <- data.frame(Depth_m = c(4, 99))
```

# Calculate contrast
```{r fig.width = 40/25.4, fig.height=50/25.4}
abun_depth_contrast <- abun_depth_mod_final |> 
  emmeans(~ Depth_m, at = depth_newdata2, type = "link") |> 
  pairs(reverse = FALSE) |> 
  gather_emmeans_draws() |> 
  summarise(median_hdci(exp(.value), .width = c(0.8, 0.95)),
            Pl = mean(.value < 0),
            Pg = mean(.value > 0)) |> 
  mutate(Contrast = "Abundance")
```

# Plot
```{r fig.width = 60/25.4, fig.height=35/25.4}
depth_abun_full_contrast <- 
ggplot() +
  geom_pointrange(aes(y = y, ymin = ymin, ymax = ymax, x = contrast), data = abun_depth_contrast |> filter(.width == 0.8), linewidth = 1.7) +
  geom_pointrange(aes(y = y, ymin = ymin, ymax = ymax, x = contrast), data = abun_depth_contrast |> filter(.width == 0.95), pch = 21, size = 1.2, fill = "#0066cc", linewidth = 0.7) +
  geom_hline(yintercept = 1, lty = 11) +
  theme_bjc() +
  scale_y_continuous("Contrast ratio", position = "right") +
  scale_x_discrete("Depth contrast 4-99m") +
   theme(axis.title = element_text(colour = "black", size = 10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(colour = "black", size = 9),
        legend.position = "none") +
    theme(plot.background = element_rect(color = "black", fill = NA, linewidth = 0.3))


depth_abun_full_contrast
```

# Save
```{r}
ggsave(filename = "output/depth_abun_full_contrast.pdf", plot = depth_abun_full_contrast, width = 50, height = 55, device = cairo_pdf, units = "mm", dpi = 300)
```


# Make RHS of plot
```{r fig.width = 60/25.4, fig.height=70/25.4}
depth_abun_fig_rhs <- 
  cowplot::plot_grid(depth_abun_effect, depth_abun_full_contrast, 
                     nrow = 2, labels = c("b", "c"), rel_heights = c(1, 1), label_size = 10) +
  theme(panel.border = element_rect(color = "black", fill = NA, linewidth = 0.3))
depth_abun_fig_rhs
```

# Make full figure

```{r fig.width = 160/25.4, fig.height=80/25.4}
depth_abun_fig_full <- 
  cowplot::plot_grid(depth_abun_trend, depth_abun_fig_rhs, 
                     nrow = 1, labels = c("a"), rel_widths = c(1, 0.6), label_size = 10) +
  theme(panel.border = element_rect(color = "black", fill = "transparent", linewidth = 0.7))
   # plot.background = element_rect(color = "black"))
depth_abun_fig_full
```

# Save
```{r}
ggsave(filename = "output/depth_abun_fig_full.pdf", plot = depth_abun_fig_full, width = 160, height = 80, device = cairo_pdf, units = "mm", dpi = 300)
ggsave(filename = "output/depth_abun_fig_full.jpg", plot = depth_abun_fig_full, width = 160, height = 80, device = "jpeg", units = "mm", dpi = 300)
```




