---
title: "Abundance modelling - Lihou, Osprey and Bougainville (LOB) - Depth as continuous predictor"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```



# Load lob_predsum data
```{r message=FALSE, warning=FALSE, include=FALSE}
source("code/packages.R")
source("code/helper_functions.R")
#sourceRmd("code/abundance/diw_abund.Rmd")
load("data/Rdata/lob_predsum.Rdata")
```

# Format of data:
lob_predsum is a tibble with each row = 1 transect of ROV predatory fish observations, summarised for total preds observed (No_Preds, i.e. abundance)


## Zero observations
```{r}
zeros <- lob_predsum %>% filter(No_Preds == 0)
```



## Can visualise the distribution by "bouncing" the zeros off the bottom of a histogram to make sense of this
```{r fig.width=140/25.4}
lob_predsum |>  
  mutate(is_zero = No_Preds == 0) |> 
  mutate(No_Preds = ifelse(is_zero, -0.1, No_Preds)) |> 
  ggplot(aes(x = No_Preds)) +
  geom_histogram(aes(fill = is_zero), binwidth = 1, boundary = 0, color = "white") +
  geom_vline(xintercept = 0) +
  scale_x_continuous() +
  scale_fill_manual(values = c("blue", "red")) +
  labs(x = "Observations of predators", 
       y = "Count",
       subtitle = "Exponential distribution with quite a few zeros") +
  theme_minimal() +
  theme(legend.position = "none")

```


# Zero inflation

We know from modelling the wider dataset that we probably will need a ZI negbinomial model. Here we are just interested in depth differences for these three reefs

## Set model formula - reef as R.E.
```{r}
abun_depth_cont_form <- bf(No_Preds ~ scale(Depth_m) + (1|Reef_1), zi ~ 1, family = zero_inflated_negbinomial(link = "log"))
```

## Check what priors we need to estimate
```{r}
get_prior(abun_depth_cont_form, data = lob_predsum)
```


# Look at range of data
```{r}
lob_predsum |> 
  group_by(Depth) |> 
  summarise(logmedian = log(median(No_Preds)), logMAD = log(mad(No_Preds)), logmean = log(mean(No_Preds)), logsd = log(sd(No_Preds)))
```



## Need a prior for the slopes (b) - weakly informative:
```{r}
abun_depth_mod_cont_priors <-  
  prior(normal(1, 2), class = 'Intercept') +
  prior(normal(0, 0.5), class = 'b') +
  prior(gamma(2, 2), class = 'shape') +
  prior(logistic(0, 1), class = 'Intercept', dpar = "zi")
```


## Fit model:
```{r}
abun_depth_mod_cont_priorsonly <- brm(abun_depth_cont_form,
  data = lob_predsum,
  prior = abun_depth_mod_cont_priors,
  sample_prior = 'only',              
  chains = 3,
  cores = 3,
  iter = 5000, 
  warmup = 1000, 
  thin = 4,
  seed = 123,
  silent = 0,
  refresh = 0,
  backend = "cmdstanr")
```


## Check how wide priors are
```{r}
abun_depth_mod_cont_priorsonly %>% conditional_effects() %>% 
   plot(points = TRUE, ask = FALSE, plot = TRUE) 
```



# Fit model with data
```{r}
abun_depth_mod_cont <- abun_depth_mod_cont_priorsonly |> update(sample_prior = 'yes', control = list(adapt_delta = 0.99, max_treedepth = 20), cores = 3, seed = 1234, refresh = 0)
```


# Check prior - posterior influence
```{r}
abun_depth_mod_cont |> SUYR_prior_and_posterior()
```


Looking fine


# Chain diagnostics

## Trace plots
```{r traceplot}
abun_depth_mod_cont$fit |>  stan_trace()
```
## Autocorrelation
```{r}
abun_depth_mod_cont$fit |> stan_ac() 
```
## rhat
```{r}
abun_depth_mod_cont$fit |> stan_rhat() 
```
## Effectiveness of sampling
```{r}
abun_depth_mod_cont$fit |> stan_ess()
```


# Residuals
```{r}
resids <- make_brms_dharma_res(abun_depth_mod_cont, integerResponse = FALSE)
testUniformity(resids)
plotResiduals(resids, form = factor(rep(1, nrow(lob_predsum))))
plotResiduals(resids, quantreg = FALSE)
testDispersion(resids) # Dispersion plot (dispersion = variance/mean) - we assume mean and variance are the same, so we need to check
```


# Residuals look ok

## PP checks
```{r}
pp_check(abun_depth_mod_cont, ndraws = 250)
```

# Logged
```{r}
predicted <- posterior_predict(abun_depth_mod_cont)
bayesplot::ppc_dens_overlay(y = log1p(lob_predsum$No_Preds), 
                            yrep = log1p(predicted[1:100,]))
```

OK so the model with Reef as random effect looks like it will work.
Howver, there may be unexplained variation as a result of "site" (e.g. could be more or less exposed, or something to do with habitat complexity etc)
So will check this against a model with site as random effect.

## Set model formula - site as re
```{r}
abun_depth_cont_form2 <- bf(No_Preds ~ scale(Depth_m) + (1|Site), zi ~ 1, family = zero_inflated_negbinomial(link = "log"))
```


# Fit model with data
```{r}
abun_depth_mod_cont2 <- 
  brm(abun_depth_cont_form2,
  data = lob_predsum,
  prior = abun_depth_mod_cont_priors,
  sample_prior = 'yes',  
  control = list(adapt_delta = 0.99, max_treedepth = 20), 
  chains = 3,
  cores = 3,
  iter = 5000, 
  warmup = 1000, 
  thin = 4,
  seed = 1234,
  silent = 0,
  refresh = 0,
  backend = "cmdstanr")
```


## Model diagnostics/validation


# Check prior - posterior influence
```{r}
abun_depth_mod_cont2 |> SUYR_prior_and_posterior()
```


Looking fine


# Chain diagnostics

## Trace plots
```{r traceplot}
abun_depth_mod_cont2$fit |>  stan_trace()
```
## Autocorrelation
```{r}
abun_depth_mod_cont2$fit |> stan_ac() 
```
## rhat
```{r}
abun_depth_mod_cont2$fit |> stan_rhat() 
```
## Effectiveness of sampling
```{r}
abun_depth_mod_cont2$fit |> stan_ess()
```


# Residuals
```{r}
resids <- make_brms_dharma_res(abun_depth_mod_cont2, integerResponse = FALSE)
testUniformity(resids)
plotResiduals(resids, form = factor(rep(1, nrow(lob_predsum))))
plotResiduals(resids, quantreg = FALSE)
testDispersion(resids) # Dispersion plot (dispersion = variance/mean) - we assume mean and variance are the same, so we need to check
```


# Residuals look ok

## PP checks
```{r}
pp_check(abun_depth_mod_cont2, ndraws = 250)
```

# Logged
```{r}
predicted <- posterior_predict(abun_depth_mod_cont2)
bayesplot::ppc_dens_overlay(y = log1p(lob_predsum$No_Preds), 
                            yrep = log1p(predicted[1:100,]))
```


The new model looks better, however there may also be unexplained variance as a result of depth. e.g. as light attenuates, perhaps we are making fewer observations due to image quality etc. We can't put depth_m (continuous variable, already fixed effect) in the model but we could use the depth_bin variable as a random intercept. Each site depth combination acts like an individual site anyway so will try this and compare with loo


## Set model formula - reef/depth as R.E.
```{r}
abun_depth_cont_form3 <- bf(No_Preds ~ scale(Depth_m) + (1|Site/Depth), zi ~ 1, family = zero_inflated_negbinomial(link = "log"))
```


# Fit model with data
```{r}
abun_depth_mod_cont3 <- 
  brm(abun_depth_cont_form3,
  data = lob_predsum,
  prior = abun_depth_mod_cont_priors,
  sample_prior = 'yes',  
  control = list(adapt_delta = 0.99, max_treedepth = 20), 
  chains = 3,
  cores = 3,
  iter = 5000, 
  warmup = 1000, 
  thin = 4,
  seed = 1234,
  silent = 0,
  refresh = 0,
  backend = "cmdstanr")
```

## Compare models with LOO compare
```{r}
loo_compare(brms::loo(abun_depth_mod_cont), # Reef as R.E.
            brms::loo(abun_depth_mod_cont2), # Site as R.E.
            brms::loo(abun_depth_mod_cont3)) # Site/Depth as R.E.
```
Model with site and depth zone as R.E. have best looic, however the S.E. in elpd diff is greater than the elpd diff itself so will use model 2 going forward


# Conditional effects plot
```{r}
#abun_depth_mod_cont |> conditional_effects() |> plot(points = FALSE, ask = FALSE) # Reef as R.E.
abun_depth_mod_cont2 |> conditional_effects() |> plot(points = FALSE, ask = FALSE) # Site as R.E.
#abun_depth_mod_cont3 |> conditional_effects() |> plot(points = FALSE, ask = FALSE) # Site/Depth as R.E.
```


# Check R2
If interested - from AIC and conditional effects plots, can predict how this is going to be:
```{r}
#bayes_R2(abun_depth_mod_cont)
bayes_R2(abun_depth_mod_cont2)
#bayes_R2(abun_depth_mod_cont3)
```



```{r}
abun_depth_mod_final <-abun_depth_mod_cont2
```

```{r}
save(abun_depth_mod_final, file = "data/Rdata/mod_data/abun_depth_mod_final.Rdata")
```


```{r}
# END #
```

