---
title: "Predator assemblage composition - species correlation in gllvm"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```


# Prerequisites
```{r message=FALSE, warning=FALSE, include=FALSE}
source("code/packages.R")
source("code/helper_functions.R")
source("code/extras.R")
load("data/Rdata/predsum.Rdata")
load("data/Rdata/preds.Rdata")
load("data/Rdata/rmetadata.Rdata")
#source("code/diversity/depth_site_diversity.R")
library(gllvm)
```

## Store list of predictors for model (mainly interested in Region for this analysis)
```{r}
env_vars_gllvm <- 
  preds |> 
  select(Region, Reef_1, Site, 
         Depth, Site_Depth,
         dCT, dGBR
         ) |> 
  distinct() |> 
  as.data.frame()
```


## Summarise species accounts by depth bin and site:
```{r}
site_spec_gllvm <- 
  preds |> 
  group_by(Site_Depth, Binomial) |> # Just the grouping variable (single) that we are interested in here
  dplyr::summarise(Count=sum(Number)) |> 
  ungroup()
```


## Convert into "wide" format diversity matrix
```{r}
site_matrix_gllvm <- 
  site_spec_gllvm |> 
  pivot_wider(names_from = Binomial, values_from=Count, values_fill = 0) |> 
  dplyr::select(!`NA`) |>  # Get rid of column where the sp name ('Binomial') had been NA
  ungroup() 

## And into mvabund object (works well with gllvm)
predabund <- mvabund(site_matrix_gllvm[2:102]) 
```



# For this analysis we are interested in assemblage a) differences between region and also b) seeing if there are identifiable latent variables behind these and c) if any species occurance correlates
# We don't know how many latent variables will best explain our data so we will try a few:
```{r}
criteria <- NULL # Create empty object to receive candidate models
for(i in 1:5){
  fiti <- gllvm(predabund, env_vars_gllvm, family = "negative.binomial", num.lv = i, sd.errors = FALSE,
                formula = ~ Region, seed = 1234)
  criteria[i] <- summary(fiti)$AICc
  names(criteria)[i] = i
}
criteria
```

# So two latent variables best model

# Run as specific model
```{r}
lvmod_region <- gllvm(predabund, env_vars_gllvm, family = "negative.binomial", num.lv = 2, sd.errors = TRUE,
                 formula = ~ Region, seed = 1234)
```


# Check residuals

```{r}
plot(lvmod_region, which = 1)
plot(lvmod_region, which = 2)
plot(lvmod_region, which = 3)
plot(lvmod_region, which = 4)
plot(lvmod_region, which = 5)
```


# Look reasonable - can't see any patterns

# Check summary - not that useful
```{r}
summary(lvmod_region)
betas <- lvmod_region$TMBfn$par %>% as.matrix()
```


```{r}
ordiplot(lvmod_region)
coefplot(lvmod_region)
```




# gllvm is useful for finding any correlations between species occurrence

## Extra libraries
```{r}
library(corrplot)
library(gclus)
```


```{r}
cr <- getResidualCor(lvmod_region)
```



# In built correlation matrix - looks horrible
```{r}
region_sp_corplot <- corrplot(cr[order.single(cr), order.single(cr)], diag = FALSE, type = "lower", 
         method = "square", tl.cex = 0.5, tl.srt = 45, tl.col = "red")
```


# But we can extract the cor scores for plotting in ggplot
```{r}
sp_cors <- tibble(region_sp_corplot$corrPos) |> 
  mutate_if(is.character, as.factor) |> 
  mutate(xName = fct_reorder(xName, x),          # Need to reorder the sp names according to the numeric position
         yName = fct_reorder(yName, y))          # For both axes
```


# If you need as a matrix
```{r}
sp_cor_matrix <- 
  sp_cors %>% 
  pivot_wider(names_from = yName, values_from = corr, id_cols = xName)
```


# We have 5460 combinations of possible correlations between our taxa and most of them are not very correlated
## Filter out the taxa with the highest negative and positive correlations
```{r}
hicor_pos <- sp_cors %>% 
  #slice_max(corr, n = 20) |> 
  filter(corr > 0.8) |> 
  rename(hicor = corr)

hicor_neg <- sp_cors %>% 
  #slice_min(corr, n = 20) |> 
  filter(corr < -0.8) |> 
  rename(hicor = corr)
```


# Combine
```{r}
sp_hicors <- bind_rows(hicor_neg, hicor_pos)
```


# And add in to sp_cors
```{r}
sp_cors1 <- 
  sp_cors |> 
  left_join(sp_hicors) |> 
  mutate(xName1 = factor(if_else(is.na(hicor), "", xName), ordered = TRUE),
         yName1 = factor(if_else(is.na(hicor), "", yName), ordered = TRUE)) |> 
  mutate(xName1 = fct_reorder(xName1, x),          # Need to reorder the sp names according to the numeric position
         yName1 = fct_reorder(yName1, y)) 
```


  
  
####

```{r}
ggplot(data = sp_cors1) +
  geom_tile(aes(x = xName, y = yName, fill = corr)) +
  #geom_tile(aes(x = xName1, y = yName1), fill = 'transparent') +
  scale_fill_gradient2("Correlation", midpoint = 0, mid ="white", limits = c(-1, +1)) + 
  #annotate()
  annotate("text", x = -1, y = 67, label = "L. gibbus", fontface = "italic", hjust = 2) +
  coord_cartesian(clip ="off") +
  scale_y_discrete(breaks = sp_cors1$yName1, labels = sp_cors1$yName1) +
  scale_x_discrete(expand = c(0, 0)) +
  theme_bw() +
  theme(plot.title = element_blank(), 
        axis.title = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        #axis.text.y = element_blank(),
        plot.margin = unit(c(1, 0, 0, 1), "cm"),           # Top, right, bottom, left
        legend.title = element_text(colour="black", size = 9)) 

```




  






























