---
title: "Predator assemblage composition using boral - Depth patterns on LOB @ transect level"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---



```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```


# Prerequisites
```{r message=FALSE, warning=FALSE, include=FALSE}
source("code/packages.R")
source("code/extras.R")
#load("data/Rdata/lob_preds.Rdata")
#load("data/Rdata/lob_predsum.Rdata")
```


# RUN fam_abund.Rmd first to get data #

# Set mcmc control parameters
```{r}
mcmc.control. <- list(n.burnin = 1000,
                      n.iteration = 7000,
                      n.thin = 30,
                      seed = 123)
```



Construct a latent variable only model, with no env predictors involved, which will allow us to construct an ordination
gllvm suggested one lv most parsimonious for these data but won't allow ordination so will go with 2
```{r}
boral_fam_LV <- boral(sitedepth_famabund_1, family="negative.binomial",
                   mcmc.control = mcmc.control.,
                   lv.control = list(num.lv = 2))
#save(boral_fam_LV, file = "data/Rdata/mod_data/boral_fam_LV.Rdata") # 
#load(file = "data/Rdata/mod_data/boral_fam_LV.Rdata") # 
```

# Model diagnostics

# Residuals
```{r}
plot(boral_fam_LV)
```

# Chain convergence 
Via geweke diagnostics
```{r}
# Raw geweke diagnostics
boral_fam_LV$geweke.diag
#geweke.plot(boral_fam_LV$jags.model)
#geweke.diag(get.mcmcsamples(boral_fam_LV))
#get.mcmcsamples(boral_fam_LV) %>% geweke.plot()
#
# Adjusted for multiple comparisons
#gew.pvals <- 2*pnorm(abs(unlist(boral_fam_LV$geweke.diag[[1]])), lower.tail = FALSE)
#p.adjust(gew.pvals, method = "holm")
```


# Model investigation - doesn't really have anything in here as it is just the latent variables behind this model fit (unconstrained)
```{r}
#tidyboral(boral_fam_LV)
#attributes(boral_fam_LV)
```


# Can also "hack" ggboral to extract site and species scores:


First need to make the plots:
```{r}
library(ggboral)
sites_plot_LV <- gg_lvsplot(boral_fam_LV, include = "objects")
species_plot_LV <- gg_lvsplot(boral_fam_LV, include = "attributes")
```

# Compare to native plot
```{r}
species_plot_LV
sites_plot_LV
lvsplot(boral_fam_LV, cex = 0.8, biplot = FALSE)
```


# Extract scores
Species scores
```{r}
spec_scores_LV <- species_plot_LV$data

# If you want the scores from the mvabund model (don't want them in this case as there aren't any!)
#sig_spec_scores_LV <- species_plot_LV$data |>  
#  rename(Binomial = label)  |> 
#  left_join(depth_TID_uni_contrasts) |>  # Need to run the mvabund Rmd first
#  filter(P <= 0.05)
```

Site scores
```{r}
site_scores_LV <- sites_plot_LV$data %>% 
  bind_cols(env_vars_fam_1)
```


# Unconstrained ordination
```{r fig.width = 100/25.4, fig.height=70/25.4}
ggplot(site_scores_LV, aes(x = lv1, y = lv2)) +
  geom_point(data = spec_scores_LV, pch = 23, fill = "orange") +
  #stat_ellipse(aes(fill = Depth, group = Depth), geom = "polygon", type = "norm" , level = 0.95, alpha = 0.6) +
  ggforce::geom_mark_hull(aes(colour = Depth, fill = Depth), expand = 0.01, concavity = 5, alpha = 0.8) + 
  geom_point(aes(fill = Depth, shape = Depth)) +
  #geom_text(data=spec_scores_LV, aes(label = label), show.legend = FALSE) +
  scale_fill_manual(values = c("lightskyblue", "dodgerblue3", "midnightblue")) +
  scale_colour_manual(values = c("lightskyblue", "dodgerblue3", "midnightblue")) +
  scale_shape_manual(values = c(24,22,25)) +
  theme_minimal()
  facet_wrap(~ Depth)

```



```{r fig.width=100/25.4, fig.height=80/25.4}
#lob_boral_ord_depth <- 
ggplot(site_scores_LV, aes(x = lv1, y = lv2)) +
  geom_segment(data=spec_scores_LV, aes(y=0, x=0, yend=lv2, xend=lv1),  color='black') +
  stat_ellipse(aes(fill = Depth, group = Depth), colour = "black", size = 0.2, geom = "polygon", level = 0.95, alpha = 0.6) +
  geom_point(aes(fill = Depth), pch = 21, size = 2, alpha = 0.5) +
  geom_point(data=spec_scores_LV, fill = "orange", pch = 24, size = 2) +
  #geom_path(data = circle, aes(y=PC2,x=PC1), color = 'black', fill = "white", linewidth = 1/2, alpha = 1/3) +
  geom_text(data=spec_scores_LV, aes(label = label), show.legend = FALSE) +
  #stat_ellipse(aes(colour = Depth, group = Depth), fill = "white", geom = "polygon", type = "norm" , level = 0.95, alpha = 1) +
  geom_hline(yintercept=0, linetype='dotted') +
  geom_vline(xintercept=0, linetype='dotted') +  
  scale_fill_manual(values = c("lightblue1", "dodgerblue3", "midnightblue")) +
  scale_colour_manual(values = c("lightblue1", "dodgerblue3", "midnightblue")) +  
  scale_shape_manual(values = c(21,22,24,25)) +
  #scale_y_continuous("L.V.2") +
 # scale_x_continuous("L.V.1", limits = c(-2.5, 2.5)) +
  theme_bjc() +
  theme(panel.background = element_rect(fill = "transparent", colour = "black"),
        legend.position = "right") 
  
#lob_boral_ord_depth
```

# Save

```{r}
ggsave(plot = lob_boral_ord_depth, filename = "output/lob_boral_ord_depth.jpg", height = 110, width = 180, device = "jpeg", units = "mm", dpi = 300)
```




# Now with depth as a predictor ("constrained"?)

```{r}

#boral_fam_depth <- boral(sitedepth_famabund_1, X = env_vars_fam_1, family="negative.binomial",
#                   mcmc.control = mcmc.control.,
#                   formula.X = ~ Depth,
#                   row.eff = "random",
#                   lv.control = list(num.lv = 2),
 #                  save.model = TRUE)

#save(boral_fam_depth, file = "data/Rdata/mod_data/boral_fam_depth.Rdata") # 
#load(file = "data/Rdata/mod_data/boral_fam_depth.Rdata") 
```


# Can also "hack" ggboral to extract site and species scores:


First need to make the plots:
```{r}
sites_plot_depth <- gg_lvsplot(boral_fam_depth, include = "objects")
species_plot_depth <- gg_lvsplot(boral_fam_depth, include = "attributes")
```

# Compare to native plot
```{r}
species_plot_depth
sites_plot_depth
lvsplot(boral_fam_LV, cex = 0.8, biplot = FALSE)
```


# Extract scores

```{r}
# Species scores
spec_scores_depth <- species_plot_depth$data

# Sites
site_scores_depth <- sites_plot_depth$data %>% 
  bind_cols(env_vars_fam_1)
```




# Basic ordination by depth
```{r fig.width = 100/25.4, fig.height=70/25.4}
ggplot(site_scores_depth, aes(x = lv1, y = lv2)) +
  geom_point(data = spec_scores_depth, pch = 23, fill = "orange") +
  geom_text(data = spec_scores_depth, mapping = aes(label = label)) +
  stat_ellipse(aes(fill = Depth, group = Depth), geom = "polygon", type = "norm" , level = 0.95, alpha = 0.6) +
  #ggforce::geom_mark_hull(aes(fill = Depth), colour = "midnightblue", expand = 0.02, concavity = 5, alpha = 0.8) + 
  #geom_point(aes(fill = Depth, shape = Depth)) +
  scale_fill_manual(values = c("lightskyblue", "dodgerblue3", "midnightblue")) +
  scale_colour_manual(values = c("lightskyblue", "dodgerblue3", "midnightblue")) +
  scale_shape_manual(values = c(24,22,25)) +
  theme_minimal()
  facet_wrap(~ Depth)

```





