---
title: "Predator assemblage composition using boral - Depth patterns on LOB @ transect level"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---



```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```


# Prerequisites
```{r message=FALSE, warning=FALSE, include=FALSE}
source("code/packages.R")
source("code/extras.R")
load("data/Rdata/lob_preds.Rdata")
load("data/Rdata/lob_predsum.Rdata")
```


# Summarising family counts

# Have to do this step separately for each reef?

Lihou
```{r}
predfams_abund_Lihou <- 
lob_preds  |>  
  filter(Reef_1 == "Lihou") |> 
  droplevels() |> 
  group_by(Reef_1, Site, Depth, Family) |> 
  summarise(Count =sum(Number, na.rm = TRUE)) |> 
  ungroup()  |> 
  filter(!is.na(Family)) |> 
  complete(Reef_1, Site, Depth, Family) %>% 
  mutate(Count = replace_na(Count, 0)) 
```

Osprey
```{r}
predfams_abund_Osprey <- 
lob_preds  |> 
  filter(Reef_1 == "Osprey") |> 
  droplevels() |>   
  group_by(Reef_1, Site, Depth, Family) |> 
  summarise(Count =sum(Number, na.rm = TRUE)) |> 
  ungroup()  |> 
  filter(!is.na(Family)) |> 
  complete(Reef_1, Site, Depth, Family) %>% 
  mutate(Count = replace_na(Count, 0)) 
```

Bougainville
```{r}
predfams_abund_Bougainville <- 
lob_preds  |>  
  filter(Reef_1 == "Bougainville") |> 
  droplevels() |>   
  group_by(Reef_1, Site, Depth, Family) |> 
  summarise(Count = sum(Number, na.rm = TRUE)) |> 
  ungroup()  |> 
  filter(!is.na(Family)) |> 
  complete(Reef_1, Site, Depth, Family) %>% 
  mutate(Count = replace_na(Count, 0)) 
```


# Bring together in 1DF - 660 combinations of site|depth ~ Family
```{r}
predfams_abund_1 <- 
  predfams_abund_Bougainville |> 
  bind_rows(predfams_abund_Osprey) |> 
  bind_rows(predfams_abund_Lihou)
```

# Diversity matrix
```{r}
sitedepth_fammatrix_1 <- 
  predfams_abund_1 |> 
  pivot_wider(names_from = Family, values_from=Count, values_fill = 0) |> 
  ungroup() 
```


# Environmental predictors
```{r}
env_vars_fam_1 <- 
  sitedepth_fammatrix_1 |> 
  select(!where(is.numeric))
```


# mvabund object - should be 17 wide (17 families)
```{r}
sitedepth_famabund_1 <- mvabund(sitedepth_fammatrix_1  |> 
                                  select(where(is.numeric)))
```


# Alt way of summarising family counts - does not back fill 0s into missing site-depth combos (3 fewer)

## Summarise family counts
```{r}
predfams_abund_2 <- 
  lob_preds |> 
  group_by(Reef_1, Site, Depth, Family) |> # Just the grouping variable (single) that we are interested in here
  dplyr::summarise(Count=sum(Number)) |> 
  ungroup()
```

## Diversity matrix
```{r}
sitedepth_fammatrix_2 <- 
  predfams_abund_2 |> 
  pivot_wider(names_from = Family, values_from=Count, values_fill = 0) |> 
  dplyr::select(!`NA`) |>  # Get rid of column where the sp name ('Binomial') had been NA
  ungroup() 
```

# Depth predictors
```{r}
env_vars_fam_2 <- 
  sitedepth_fammatrix_2 |> 
  select(!where(is.numeric))
```

# mvabund object - should be 17 wide (17 families)
```{r}
sitedepth_famabund_2 <- mvabund(sitedepth_fammatrix_2 |> 
                                  select(where(is.numeric)))
```


## Mean-variance relationship
```{r}
par()
meanvar.plot(sitedepth_famabund_1)
meanvar.plot(sitedepth_famabund_2)
```

# Model summaries
```{r}
# Neg binom model options
fam_mva_1 <- manyglm(sitedepth_famabund_1 ~ env_vars_fam_1$Depth, family="negative_binomial", )
fam_mva_2 <- manyglm(sitedepth_famabund_2 ~ env_vars_fam_2$Depth, family="negative_binomial")
```


# Model validation 
# In mvabund the only way to check this is plotting residuals vs fitted values
```{r fig.width=300/25.4, fig.height=100/25.4}
plot.manyglm(fam_mva_1, which = c(1,2)) # 
plot.manyglm(fam_mva_2, which = c(1,2)) # 
```

# Summaries
```{r}
fam_mvasum <- summary(fam_mva_1)
fam_mvasum
```


# Pairwise comparisons
```{r}
fam_mva_aov_1 <- anova.manyglm(fam_mva_1, test = 'LR', pairwise.comp = env_vars_fam_1$Depth,  p.uni="adjusted")
```


# Results

# Multivariate contrasts
```{r}
fam_multi_contrasts_1 <- 
  fam_mva_aov_1$pairwise.comp.table  |> 
              rownames() |> 
              as_tibble() |> 
              mutate(Contrast = value, .keep = "unused") |> 
  bind_cols(fam_mva_aov_1$pairwise.comp.table  |> 
  as_tibble())

fam_multi_contrasts_1
```

# Univariate contrasts
```{r}
fam_uni_contrasts_1 <- 
  t(fam_mva_aov_1$uni.test) %>% 
  as.data.frame() %>% 
  rownames_to_column("Family") %>% 
  rename(LRT = `env_vars_fam_1$Depth`) %>% 
  select(!`(Intercept)`) %>% 
  left_join(  
  t(fam_mva_aov_1$uni.p) %>% 
  as.data.frame() %>% 
  rownames_to_column("Family") %>% 
  rename(P = `env_vars_fam_1$Depth`) %>% 
  select(!`(Intercept)`)) %>% 
  arrange(-LRT)

fam_uni_contrasts_1
```



# Extras

# Predator family abundance (alt way)
```{r}
predfams_abund <- 
  lob_preds  |>  
  group_by(Reef_1, Site, Depth, Family) |> 
  summarise(No_Preds =sum(Number, na.rm = TRUE)) |> 
  ungroup()  |> 
  filter(!is.na(Family)) |> 
  complete(Reef_1, Depth, Family) %>% 
  mutate(No_Preds = replace_na(No_Preds, 0))
```


# All families abundance summary (mean ±SE) over depth categories
```{r fig.width=140/25.4, fig.height=87/25.4}
predfams_abund_sum <- 
lob_preds  |>  
  group_by(Reef_1, Depth, Family) |> 
  summarise(No_Preds =sum(Number, na.rm = TRUE)) |> 
  ungroup()  |> 
  filter(!is.na(Family)) |> 
  complete(Reef_1, Depth, Family) %>% 
  mutate(No_Preds = replace_na(No_Preds, 0)) |> 
  group_by(Depth, Family) %>%
  summarise(mean = mean(No_Preds, na.rm = TRUE),
            sd = sd(No_Preds, na.rm = TRUE),
            SE = std.error(No_Preds, na.rm = TRUE))  |> 
  mutate(Family = fct_reorder(Family, -mean)) |> 
  arrange(Family) |> 
  ungroup()
```


# Plot - wide
```{r fig.width=160/25.4, fig.height=90/25.4}
predfams_abund_fig <- 
predfams_abund_sum  |> 
  ggplot(aes(x = Depth, y = mean, group = Family)) +
  geom_line(aes()) +
  geom_errorbar(aes(ymin = mean - SE, ymax = mean + SE), width = 0.15, size = 0.4, colour = "black", position = position_dodge(width = 1), show.legend = FALSE) +
  geom_point(aes(shape = Depth, fill = Depth), colour = "black", size = 2, position = position_dodge(width = 1), show.legend = TRUE) +
  scale_y_continuous(name = expression("Abundance (n.150m"^{2}~" ± SE)"), expand = c(0.1, 0.1)) +
  scale_fill_manual(name = "Depth zone", values = c("lightskyblue", "dodgerblue3", "midnightblue"), 
                    labels = c("Shallow\n(0-30m)", "Upper\n(31-60m)", "Lower\n(61-100m)")) +
  scale_shape_manual(name = "Depth zone", values = c(24, 23, 25), 
                     labels = c("Shallow\n(0-30m)", "Upper\n(31-60m)", "Lower\n(61-100m)")) +
  scale_x_discrete(name = "Family") +
  theme_bjc() + 
  theme(legend.position = c(0.68, 0.06),
        #legend.title = element_blank(),
        legend.text = element_text(),
        legend.direction = "horizontal",
        axis.title.x = element_blank(),
        strip.text = element_text(size = 9, family = "Arial", colour = "black"),
        axis.text.x = element_blank(),
        plot.background = element_rect(fill = "transparent", colour = "black")) +
   guides(shape = guide_legend(override.aes = list(size = 4))) +
   guides(fill = guide_legend(title.position = "top", title.hjust=0.5)) +
   facet_wrap(~Family, scales = "free_y", strip.position = "bottom")

predfams_abund_fig
```
# Save
```{r}
ggsave(filename = "output/family_depth_abun_fig.jpg", plot = predfams_abund_fig, width = 160, height = 99, device = "jpeg", units = "mm", dpi = 300)
```


# Plot - long
```{r fig.width=110/25.4, fig.height=160/25.4}
predfams_abund_fig_long <- 
predfams_abund_sum  |> 
  ggplot(aes(x = Depth, y = mean, group = Family)) +
  geom_line(aes()) +
  geom_errorbar(aes(ymin = mean - SE, ymax = mean + SE), width = 0.15, size = 0.4, colour = "black", position = position_dodge(width = 1), show.legend = FALSE) +
  geom_point(aes(shape = Depth, fill = Depth), colour = "black", size = 2, position = position_dodge(width = 1), show.legend = TRUE) +
  scale_y_continuous(name = expression("Abundance (n.150m"^{2}~" ± SE)"), expand = c(0.1, 0.1)) +
  scale_fill_manual(name = "Depth zone", values = c("lightskyblue", "dodgerblue3", "midnightblue"), 
                    labels = c("Shallow (0-30m)", "Upper (31-60m)", "Lower (61-100m)")) +
  scale_shape_manual(name = "Depth zone", values = c(24, 23, 25), 
                     labels = c("Shallow (0-30m)", "Upper (31-60m)", "Lower (61-100m)")) +
  scale_x_discrete(name = "Family") +
  theme_bjc() + 
  theme(legend.position = c(0.82, 0.06),
        legend.title = element_blank(),
        legend.text = element_text(),
        legend.direction = "vertical",
        axis.title.x = element_blank(),
        strip.text = element_text(size = 9, family = "Arial", colour = "black"),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 9, family = "Arial", colour = "black"),
        plot.background = element_rect(fill = "transparent", colour = "black")) +
   guides(shape = guide_legend(override.aes = list(size = 3))) +
   guides(fill = guide_legend(title.position = "top", title.hjust=0.5)) +
   facet_wrap(~Family, scales = "free_y", strip.position = "bottom", nrow = 6)


predfams_abund_fig_long
```

# Save
```{r}
ggsave(filename = "output/family_depth_abun_fig_long.jpg", plot = predfams_abund_fig_long, height = 160, width = 110, device = "jpeg", units = "mm", dpi = 300)
```






# As bar plots
## Separate into two levels first - the "abundant" vs "rare"
We'll call them top vs bottom

```{r}
top_predfams <- 
lob_preds  |>  
  group_by(Reef_1, Depth, Family) |> 
  summarise(No_Preds =sum(Number, na.rm = TRUE)) |>  # All good - 514
  ungroup()  |> 
  filter(!is.na(Family)) |> 
  complete(Reef_1, Depth, Family) %>% 
  mutate(No_Preds = replace_na(No_Preds, 0)) |> 
  group_by(Depth, Family) %>%
  summarise(mean = mean(No_Preds, na.rm = TRUE),
            sd = sd(No_Preds, na.rm = TRUE),
            SE = std.error(No_Preds, na.rm = TRUE))  |> 
  mutate(Family = fct_reorder(Family, -mean)) |> 
  arrange(Family) %>% 
  slice_head(n = 9)
```


```{r fig.width=140/25.4, fig.height=87/25.4}
bottom_predfams <- 
lob_preds  |>  
  group_by(Reef_1, Depth, Family) |> 
  summarise(No_Preds =sum(Number, na.rm = TRUE)) |>  # All good - 514
  ungroup()  |> 
  filter(!is.na(Family)) |> 
  complete(Reef_1, Depth, Family) %>% 
  mutate(No_Preds = replace_na(No_Preds, 0)) |> 
  group_by(Depth, Family) %>%
  summarise(mean = mean(No_Preds, na.rm = TRUE),
            sd = sd(No_Preds, na.rm = TRUE),
            SE = std.error(No_Preds, na.rm = TRUE))  |> 
  mutate(Family = fct_reorder(Family, -mean)) |> 
  arrange(Family) %>% 
  slice_tail(n = 8)
```


```{r fig.width=160/25.4, fig.height=60/25.4}
top_predfams_plot <- 
  top_predfams %>% 
  ggplot(aes(x = Family, y = mean)) +
  geom_col(aes(fill = Depth, group = Depth), colour = "black", width = 0.7, position = position_dodge(width = 0.8)) +
  geom_linerange(aes(ymin = mean - SE, ymax = mean + SE, group = Depth), size = 0.4, colour = "black", position = position_dodge(width = 0.8)) +
#  scale_y_continuous(name = expression("Abundance (n.150m"^{2}~" ± SE)"), expand = c(0.1, 0.1)) +
  scale_y_continuous(name = "\n", expand = c(0.1, 0.1)) +
  scale_fill_manual(values = c("lightskyblue", "dodgerblue3", "darkblue")) +
  scale_x_discrete(name = "Family") +
  theme_bjc() + 
  theme(legend.position = "none",
        legend.title = element_blank(),
        axis.title.x = element_blank(),
        #plot.background = element_rect(fill = "transparent", colour = "black"),
        #strip.text = element_text(size = 10, face = "italic"),
        axis.text.x = element_text(size = 10, face = "plain", angle = 45, vjust = 0.75))
top_predfams_plot
```


```{r fig.width=160/25.4, fig.height=50/25.4}
bottom_predfams_plot <- 
  bottom_predfams  |> 
  ggplot(aes(x = Family, y = mean)) +
  geom_col(aes(fill = Depth, group = Depth), colour = "black", width = 0.7, position = position_dodge(width = 0.8)) +
  geom_linerange(aes(ymin = mean - SE, ymax = mean + SE, group = Depth), size = 0.4, colour = "black", position = position_dodge(width = 0.8)) +
#  scale_y_continuous(name = expression("Abundance (n.150m"^{2}~" ± SE)"), expand = c(0.1, 0.1)) +
  scale_y_continuous(name = "\n", expand = c(0.1, 0.1)) +
  scale_fill_manual(values = c("lightskyblue", "dodgerblue3", "darkblue")) +
  scale_x_discrete(name = "Family") +
  theme_bjc() + 
  theme(legend.position = "none",
        legend.title = element_blank(),
        axis.title.x = element_blank(),
        #plot.background = element_rect(fill = "transparent", colour = "black"),
        axis.text.x = element_text(size = 10, face = "plain", angle = 45, vjust = 0.75))
bottom_predfams_plot
```



```{r fig.width=160/25.4, fig.height=100/25.4}
cowplot::plot_grid(top_predfams_plot, bottom_predfams_plot, nrow = 2) +
  annotate("text", x = 0.02, y = 0.5, label = expression("Abundance (n.150m"^{2}~" ± SE)"), angle = 90) +
  theme(plot.background = element_rect(fill = "transparent", colour = "black"))
```


















# JUNK #


# All families - 10m depth bin
```{r fig.width=160/25.4, fig.height=180/25.4}
lob_preds  |>  
  group_by(Reef_1, Depth_bin, Family) |> 
  summarise(No_Preds =sum(Number), na.rm = TRUE) |>  # All good - 514
  ungroup()  |> 
  complete(Reef_1, Depth_bin, Family) %>% 
  mutate(No_Preds = replace_na(No_Preds, 0)) |> 
  group_by(Depth_bin, Family) %>%
  summarise(mean = mean(No_Preds),
            sd = sd(No_Preds, na.rm = TRUE),
            SE = std.error(No_Preds, na.rm = TRUE)) %>% 
  ggplot(aes(x = Depth_bin, y = mean, fill = Depth_bin)) +
  geom_line(aes(group = 1)) +
  geom_linerange(aes(ymin = mean - SE, ymax = mean + SE)) +
  geom_point(size = 1) +
  scale_y_continuous(name = expression("Mean abundance (.150m"^{2}~" ± SE)"), expand = c(0.1, 0.1)) +
  scale_x_discrete(name = "Depth bin (m)") +
  theme_bjc() + 
  theme(legend.position = "none",
        legend.title = element_blank(),
        strip.text = element_text(size = 10, face = "italic"),
        axis.text.x = element_blank(),
        plot.background = element_rect(fill = "transparent", colour = "black")) +
  facet_wrap(~ Family, nrow = 6, scales = "free_y")

  
```


