---
title: "Predator assemblage composition using boral - Depth patterns on LOB"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---



```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```


# Prerequisites
```{r message=FALSE, warning=FALSE, include=FALSE}
source("code/packages.R")
source("code/extras.R")
load("data/Rdata/lob_preds.Rdata")
load("data/Rdata/lob_predsum.Rdata")
```



## Summarise species accounts by depth bin and site:
```{r}
depth_site_speccount <- 
  lob_preds |> 
  group_by(Region, Reef_1, Site, Depth, Site_Depth, Binomial) |> 
  dplyr::summarise(Count=sum(Number)) |> 
  ungroup()
```


## Convert into "wide" format diversity matrix - this is going to be our y variable
```{r}
depth_site_abund <- 
  depth_site_speccount |> 
  pivot_wider(names_from = Binomial, values_from=Count, values_fill = 0) |> 
  dplyr::select(!`NA`) |>  # Get rid of column where the sp name ('Binomial') had been NA
  ungroup()
```


# And convert into matrix
```{r}
depth_site_matrix <- 
  depth_site_abund |> 
  select(!c(Region, Reef_1, Site, Depth, Site_Depth)) #|> 
  as.matrix()
```

## Store list of predictors - this is going to be our x variable
```{r}
env_boral_sitedepth <- 
  lob_preds |> 
  select(Reef_1, Site,
         Depth, Site_Depth
         ) |> 
  distinct() |> 
  as.data.frame() 

env_boral_sitedepth$Depth
```


# Set mcmc control parameters
```{r}
mcmc.control. <- list(n.burnin = 1000,
                      n.iteration = 7000,
                      n.thin = 30,
                      seed = 123)
```



# Step 1.

We can construct a latent variable only model, with no env predictors involved, which will allow us to construct an ordination:
```{r}
#boral_nb3 <- boral(depth_site_matrix, family="negative.binomial",
#                   mcmc.control = mcmc.control.,
#                   lv.control = list(num.lv = 2))
#save(boral_nb3, file = "data/Rdata/mod_data/boral_nb3.Rdata") # 
load(file = "data/Rdata/mod_data/boral_nb3.Rdata") # 
```


# So we can check residuals
```{r}
plot(boral_nb3)
```


# Model investigation - doesn't really have anything in here as it is just the latent variables behind this model fit (unconstrained)
```{r}
tidyboral(boral_nb3)
attributes(boral_nb3)
```



# From this model we can make a basic (horrible) biplot:
```{r}
lvsplot(boral_nb3, cex = 0.8)
```

# Or with species names removed
```{r}
lvsplot(boral_nb3, cex = 0.8, 
        biplot = FALSE)
```



# Need to extract these values to make a decent plot

```{r}
boral_nb3$lv.coefs.median
```


# Sites
```{r}
site.coefs <- 
  env_boral_sitedepth |> 
  bind_cols(boral_nb3$lv.median %>% 
  as_tibble()) |> 
  mutate(Row_No = seq(1, 48, 1))
```


# Plot
```{r}
site.coefs %>% 
  ggplot() +
  #geom_point(aes(x = lv2, y = lv1)) +
  geom_text(aes(x = lv2, y = lv1, label = Row_No))

```


We could tidy this up for plotting but instead we'll refit with the env_vars




# Step 2 Fit with some predictor variables

```{r}
#boral_nb4 <- boral(depth_site_matrix, X = env_boral_sitedepth, family="negative.binomial",
#                   formula.X = ~ Depth,
#                   #ranef.ids = env_boral_sitedepth$Site,
#                   row.eff = "random",
#                   mcmc.control = mcmc.control.,
#                   lv.control = list(num.lv = 2),
#                   save.model = TRUE)

#save(boral_nb4, file = "data/Rdata/mod_data/boral_nb4.Rdata") # 
load(file = "data/Rdata/mod_data/boral_nb4.Rdata") # 

```


# Check geweke diagnostics (chain convergence)
```{r}
# Raw geweke diagnostics
boral_nb4$geweke.diag

# Adjusted for multiple comparisons
gew.pvals <- 2*pnorm(abs(unlist(boral_nb4$geweke.diag[[1]])), lower.tail = FALSE)
p.adjust(gew.pvals, method = "holm")
```


# Redo the biplot
Looks horrible:
```{r}
plot(boral_nb4$lv.median[,1] ~  
       boral_nb4$lv.median[,2],
     ylab = "",xlab = "",
     main= "un-rotated biplot")
```

# So we'll make our own
```{r}
x <- boral_nb4
```


```{r}
#alpha is an argument in lvsplot(); relates to scaling
alpha <- 0.5
```


```{r}
testcov <- x$lv.median %*% t(x$lv.coefs.median[, 2:3])

#singular value decom
do.svd <- svd(testcov, x$num.lv, x$num.lv)

#svd has components d, u and v

#some math related to rescaling
choose.lvs <- scale(do.svd$u *      matrix(do.svd$d[1:x$num.lv]^alpha, 
nrow = x$n, ncol = 2, byrow = T), center = T, scale = F)

#some more math related to rescaling
choose.lv.coefs <- scale(do.svd$v * matrix(do.svd$d[1:x$num.lv]^(1 - 
alpha), nrow = x$p, ncol = 2, byrow = T), center = T, 
                             scale = F)
```


# Check that our calculations produce the same scaled biplot:
```{r}
lvsplot(boral_nb4, cex = 0.8, biplot = F, main = "real biplot")
plot(choose.lvs, main = "hacked biplot")
```

Looks good, can plot properly now but will take a bunch more steps




```{r fig.width=270/25.4, fig.height=100/25.4}
#boral_ord_data <- 
  env_boral_sitedepth |> 
  bind_cols(choose.lvs  |>  as.data.frame()) |> 
  ggplot(aes(x = V1, y = V2)) +
  geom_hline(yintercept = 0, lty = 11) +
  geom_vline(xintercept = 0, lty = 11) +
  geom_point(aes(colour = Depth, shape = Depth)) +
 # geom_point(data = taxa_coefs) +
 # geom_text(aes(label = Taxon), data = taxa_coefs) +
  ggforce::geom_mark_hull(aes(fill = Depth), expand=0.02, concavity = 3) + 
  #ggforce::geom_mark_ellipse(aes(fill = Region), expand = 0, tol = 0.001) + 
  theme_bw() +
  facet_wrap(~ Depth)
  
```





```{r}
summary(boral_nb4)
```




# Extract species coefficients
```{r}
# Mean values
Xcoefs_mean <-
  boral_nb4$X.coefs.mean |> 
  as.matrix() |> 
  as.data.frame() |> 
  rownames_to_column(var = "Species")  |> 
  pivot_longer(cols = !Species, names_prefix = "Depth", names_to = "Depth", values_to = "Mean")

# Median values
Xcoefs_median <-
  boral_nb4$X.coefs.median |> 
  as.matrix() |> 
  as.data.frame() |> 
  rownames_to_column(var = "Species")  |> 
  pivot_longer(cols = !Species, names_prefix = "Depth", names_to = "Depth", values_to = "Median")
    
# Lower CI
Xcoefs_lowerHPD <- 
boral_nb4$hpdintervals$X.coefs |> 
  as.data.frame()  |> 
  rownames_to_column(var = "Species") |> 
  select(!4:5) |> 
  pivot_longer(cols = !Species, names_prefix = "Depth", names_pattern = "(.*).lower", names_to = "Depth", values_to = "lower.HPD")

# Upper CI
Xcoefs_upperHPD <- 
boral_nb4$hpdintervals$X.coefs |> 
  as.data.frame()  |> 
  rownames_to_column(var = "Species") |> 
  select(!2:3) |> 
  pivot_longer(cols = !Species, names_prefix = "Depth", names_pattern = "(.*).upper", names_to = "Depth", values_to = "upper.HPD")
```

# Combine
```{r}
Xcoefs <- 
  Xcoefs_median |> 
  full_join(Xcoefs_lowerHPD) |> 
  full_join(Xcoefs_upperHPD) |> 
  mutate_if(is.character, as.factor)
```

# Now we could use some higher taxonomy info for plotting

# Create family order/index
```{r}
family_order <- 
  lob_preds |> 
  select(Family) |> 
  distinct() |> 
  arrange(Family) |> 
  mutate(Fam_Ord = seq(1,19,1))
```

# Get family info from preds
```{r}
family_info <- 
  lob_preds |> 
  select(Family, Binomial) |> 
  distinct() |> 
  rename(Species = Binomial) |> 
  left_join(family_order) # And add the order back in 
```


# Extract just the significant differences in taxa
```{r}
sigXcoefs <- Xcoefs |> 
  left_join(family_info) |> 
      left_join(family_info) |> 
  mutate(TestA = ifelse(lower.HPD < 0, -1, 0),
         TestB = ifelse(upper.HPD > 0, 1, 0),
         Flag = TestA+TestB) |> 
  filter(Flag != 0) |> 
  mutate(Species = fct_reorder(Species, Fam_Ord))  |> 
  mutate(Depth = fct_relevel(Depth, "Upper"))
```



# What are the significant taxa?
```{r}
sigXcoefs |> 
  select(Species) %>% 
  distinct()
```

13 different species behind depth-related differences in assemblage composition

# Plot significant taxa
```{r fig.width = 160/25.4, fig.height=100/25.4}
lob_taxa_depth_diffs <- 
sigXcoefs |> 
  mutate(Depth = factor(case_when(grepl("Upper", Depth) ~ "Upper mesophotic",
                                grepl("Lower", Depth) ~ "Lower mesophotic"))) |> 
  mutate(Depth = fct_relevel(Depth, "Upper mesophotic")) |> 
  ggplot(aes(x = Species, y = Median)) +
  geom_pointrange(aes(ymin = lower.HPD, ymax = upper.HPD), colour = "grey30") +
  geom_point(aes(fill = Family), pch = 21, colour = "black", size = 3) +
  geom_hline(yintercept = 0, lty = 11) +
  scale_fill_viridis_d() +
  scale_colour_viridis_d() +
  scale_y_continuous("L.V. coef. contrasts") +
  coord_flip() +
  theme_bjc() +
  theme(axis.text.x = element_blank()) +
  facet_wrap(~ Depth) +
  theme(axis.text.y = element_text(face = 'italic'))

lob_taxa_depth_diffs
```



```{r}
ggsave(filename = "output/lob_taxa_depth_diffs.pdf", plot = lob_taxa_depth_diffs, height = 100, width = 170, device = cairo_pdf, units = "mm", dpi = 300)
```



## Next, we'll marry this up with lob_preds to demonstrate abundance with depth for each taxon

# What are the significant taxa?
```{r}
sig_taxa <- 
sigXcoefs |> 
  select(Species) %>% 
  distinct()  |> 
  mutate(Species = as.character(Species)) |> 
  arrange(Species)
sig_taxa
```

```{r}
sig_lob_preds <- 
  sig_taxa |> 
  select(Species) %>% 
  distinct() %>% 
  rename(Binomial = Species) %>% 
  left_join(lob_preds)
```


```{r}
sig_lob_predsum <- 
   sig_lob_preds  %>% 
   group_by(Depth, Reef_1, Binomial) %>%
   summarise(No_Preds =sum(Number)) %>%  # All good - 514
   ungroup() %>% 
  complete(Depth, Binomial, Reef_1) %>% 
  mutate(No_Preds = replace_na(No_Preds, 0))
```

```{r}
sig_lob_predsum2 <- sig_lob_predsum %>% 
  group_by(Depth, Binomial) %>%
  summarise(mean = mean(No_Preds),
            sd = sd(No_Preds, na.rm = TRUE),
            SE = std.error(No_Preds, na.rm = TRUE)) %>% 
  arrange(-mean) %>% 
  arrange(Depth) 
```




```{r fig.width=150/25.4, fig.height=180/25.4}
sig_abun_plot <- 
sig_lob_predsum2 %>% 
  ggplot(aes(x = Depth, y = mean, fill = Depth)) +
  geom_line(aes(group = 1)) +
  geom_linerange(aes(ymin = mean - SE, ymax = mean + SE)) +
  geom_point(pch = 21, size = 2) +
  scale_y_continuous(name = "Mean abundance ± SE") +
  theme_bjc() +
  theme(legend.title = element_blank(),
        legend.position = c(0.7, 0.07),
        legend.direction = "horizontal",
        axis.text = element_blank(),
        strip.text.x = element_text(face = "italic")) +
  facet_wrap(~ Binomial, nrow = 9, scales = "free_y")

sig_abun_plot
```









# Redo ordination with taxon labels


# Extract taxa list from model
```{r}
boral_taxa_list <- 
  colnames(boral_nb4$y) |> 
  as_tibble() |> 
  rename(Species = value)
```


# Convert the coefficient matrices into tbls:
```{r}
taxa_coefs <- choose.lv.coefs |> 
  as_tibble() |> 
  bind_cols(boral_taxa_list) |> 
  right_join(sigXcoefs[1])
```

```{r fig.width=270/25.4, fig.height=100/25.4}
#boral_ord_data <- 
  env_boral_sitedepth |> 
  bind_cols(choose.lvs  |>  as.data.frame()) |> 
  ggplot(aes(x = V1, y = V2)) +
  geom_hline(yintercept = 0, lty = 11) +
  geom_vline(xintercept = 0, lty = 11) +
  geom_point(aes(colour = Depth, shape = Depth)) +
 # geom_point(data = taxa_coefs) +
  geom_text(aes(label = Species), data = taxa_coefs) +
  ggforce::geom_mark_hull(aes(fill = Depth), expand=0.02, concavity = 3) + 
  #ggforce::geom_mark_ellipse(aes(fill = Region), expand = 0, tol = 0.001) + 
  theme_bw() +
  facet_wrap(~ Depth)
  
```



# Compare to native biplot:
```{r}
lvsplot(boral_nb4, cex = 0.8)
```




