---
title: "Predator assemblage composition - mvabund @ site_depth level"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```


# Prerequisites
```{r message=FALSE, warning=FALSE, include=FALSE}
source("code/packages.R")
source("code/extras.R")
load("data/Rdata/lob_predsum.Rdata")
load("data/Rdata/lob_preds.Rdata")
```


## Store list of predictors we'll be interested in at the end.
```{r}
env_vars_sitedepth <- 
  lob_preds |> 
  select(Reef_1, Site, Depth, Site_Depth) |> 
  distinct() #|> 
#  right_join(transect_abun_matrix)
```



## Summarise species accounts by depth bin and site:
```{r}
sitedepth_speccount <- 
  lob_preds |> 
  group_by(Site_Depth, Binomial) |> # Just the grouping variable (single) that we are interested in here
  dplyr::summarise(Count=sum(Number)) |> 
  ungroup()
```




## Convert into "wide" format diversity matrix
```{r}
sitedepth_matrix <- 
  sitedepth_speccount |> 
  pivot_wider(names_from = Binomial, values_from=Count, values_fill = 0) |> 
  dplyr::select(!`NA`) |>  # Get rid of column where the sp name ('Binomial') had been NA
  ungroup() 
```


## Turn into mvabund object
```{r}
#predabund <- mvabund(site_matrix[-1]) 
predabund_SDdepth <- mvabund(sitedepth_matrix[-1]) 
```

## Check out mean-variance relationship - should be 1-1 as we are working with count data
```{r}
par(predabund_SDdepth)
meanvar <- meanvar.plot(predabund_SDdepth) 
```

# EDA - can do a basic overview of abundance ~ predictor variable
Won't work in Rmd!! Need to copy to script or console
```{r}
#plot(predabund ~ env_vars_sitedepth$Depth)
```


#  Model fitting
## Notes:
- mvabund cannot handle random factors (yet) so a resampling approach used instead (see below)
- There is a pattern in the Poisson model residuals (not surprising, given the ZI in these data) so have commented out

```{r}
# Neg binom model options
depth_SD_mod <- manyglm(predabund_SDdepth ~ env_vars_sitedepth$Depth, family="negative_binomial")
```



# Model validation 
# In mvabund the only way to check this is plotting residuals vs fitted values
```{r}
plot.manyglm(depth_SD_mod, which = 1) # 
plot.manyglm(depth_SD_mod, which = 2) # 
plot.manyglm(depth_SD_mod, which = 3) # 
```

Residuals looking ok




# Summaries
```{r}
#depth_SD_mvasum <- summary(depth_SD_mod)
#save(depth_SD_mvasum, file = "data/Rdata/mod_data/depth_SD_mvasum.Rdata") # 
load(file = "data/Rdata/mod_data/depth_SD_mvasum.Rdata") # 
depth_SD_mvasum
```


# Pairwise comparisons

```{r}
#depth_SD_aov <- anova.manyglm(depth_SD_mod, test = 'LR', pairwise.comp = env_vars_sitedepth$Depth,  p.uni="adjusted")
```

# These anovas are very slow to run -save results for use later
```{r}
#save(depth_SD_aov, file = "data/Rdata/mod_data/mva_depth_SD_aov.rda") # 
```

# Reload if/when required
```{r}
load(file = "data/Rdata/mod_data/mva_depth_SD_aov.rda")
```

# Results

# Multivariate contrasts
```{r}
depth_SD_multi_contrasts <- 
  depth_SD_aov$pairwise.comp.table  |> 
              rownames() |> 
              as_tibble() |> 
              mutate(Contrast = value, .keep = "unused") |> 
  bind_cols(depth_SD_aov$pairwise.comp.table  |> 
  as_tibble())

depth_SD_multi_contrasts
```

# Univariate contrasts
```{r}
depth_SD_uni_contrasts <- 
  t(depth_SD_aov$uni.test) %>% 
  as.data.frame() %>% 
  rownames_to_column("Binomial") %>% 
  rename(LRT = `env_vars_sitedepth$Depth`) %>% 
  select(!`(Intercept)`) %>% 
  left_join(  
  t(depth_SD_aov$uni.p) %>% 
  as.data.frame() %>% 
  rownames_to_column("Binomial") %>% 
  rename(P = `env_vars_sitedepth$Depth`) %>% 
  select(!`(Intercept)`)) %>% 
  mutate(Binomial = str_replace(Binomial, "\\.", " ")) %>% 
  arrange(-LRT)

depth_SD_uni_contrasts
```


