---
title: " Mapping LOB Overview"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    #fig_height: 
    fig_width: 140/25.4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---
## Created: 3rd Oct 2023 BJC ##



```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```



# Load required packages
```{r packages, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
source('code/packages.R')
library(sf)
library(rio)
library(ggsn)
```


# Load data

## Load Australia map object from ggplot's world map
Note: could use whole world map but the file will be 50 x smaller if we filter now (0.1 vs ~5MB), and thus quicker to plot and work with
```{r world-map-load}
aus <- map_data("world") %>% 
  filter(region == 'Australia')
```

## Load in shape files 
Need to have the .shp and the .shx file in the same folder
### CS
```{r load-shp-sf, message=FALSE, warning=FALSE}
lob_dryreef <- read_sf("data/3dgbr_geomorph/shape/coralsea_dryreef.shp") |> 
    filter(NAME == "Osprey Reef" |
           NAME == "Bougainville Reef" |
           NAME == "Lihou Reef")

lob_reef <- read_sf("data/3dgbr_geomorph/shape/coralsea_reef.shp") |> 
    filter(NAME == "Osprey Reef" |
           NAME == "Bougainville Reef" |
           NAME == "Lihou Reef")

```


### GBR
```{r load-shp-sf2, message=FALSE, warning=FALSE}
gbr_feats <- read_sf("data/3dgbr_geomorph/shape/gbr_features.shp")
gbr_shelf <- read_sf("data/3dgbr_geomorph/shape/gbr_shelf.shp") # Don't need to filter as the file is already as small as it can be
```


### Create reef labels
```{r}
reef_labels <- 
  lob_predsum %>% 
  select(Reef_1, Site, Site_lat, Site_long) %>% 
  unique() %>% 
  group_by(Reef_1) %>% 
  slice_max(Site_lat, n = 1) |> 
  select(!Site)
```


```{r}
reef_labels <- read_excel("data/reef_labels.xlsx")
```



### Town labels
```{r}
town_labels <- 
  tibble(Town = c("Cairns", "Townsville"),
         Lat = c(-16.93, -19.29), 
         Long = c(145.77, 146.77))      


```



## Basic map trimmed down to central and northern Coral Sea (i.e. just keeping in LOB)
- Could specify detailed aes here but will save that step for the final maps to keep this early code easy to follow
```{r}
australia_map <- 
  aus %>% 
  ggplot() +
  geom_sf(data = gbr_feats, fill = 'grey60', colour = "transparent", size = .001) +
  geom_polygon(aes(x = long, y = lat, group = group), fill = "grey40", colour = 'black', linewidth = 0.3) +
  coord_sf(ylim = c(-19, -10), xlim = c(142.5, 152)) 

australia_map
```

If file download required see mapping_ROV.Rmd



## LOB combined overview general area
```{r fig.width= 150/25.4, fig.height=150/25.4}
overview_map <- 
australia_map +
  geom_segment(data = reef_labels, mapping = aes(x = Site_long, xend = Label_long, y = Site_lat, yend = Label_lat), size = 0.25) +
  geom_label(data = reef_labels, mapping = aes(x = Label_long, y = Label_lat, label = Reef_1), family = "Arial", size = 4, fill = "white", colour = "black", label.size = NA) +
  geom_sf(data = lob_dryreef, fill = 'gray60', size = 0.1) +
  geom_sf(data = lob_reef, fill = 'grey', colour = 'black', size = 0.1) +
  geom_text_repel(data = town_labels, mapping = aes(x = Long, y = Lat, label = Town), nudge_x = -0.5) +
  geom_point(data = town_labels, mapping = aes(x = Long, y = Lat), pch = 21, fill = "red", colour = "black", size = 3) +
  coord_sf(ylim = c(-20, -10.2), xlim = c(142.8, 153)) +
  theme(axis.title = element_blank(), 
        axis.text = element_text(colour = "black", size = 10),
        #axis.ticks = element_blank(),
        #panel.grid = element_line(colour = "#ECEFF1"),
        panel.border = element_rect(colour = "black", fill = "transparent"),
        panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "transparent", colour = "transparent"))

overview_map
```


# Add globe and scale bar
```{r}
sourceRmd::sourceRmd("code/make_globe.Rmd")
```


```{r fig.width= 140/25.4, fig.height=140/25.4}
lob_map_midrange <- 
  overview_map +
  #annotation_custom(ggplotGrob(globe), xmin = 141, xmax = 147, ymin = -20, ymax = -17) +
  scalebar(x.min = 140, x.max = 145, y.min = -20, y.max = -19, dist = 100, dist_unit = "km", transform = TRUE, model = "WGS84", # Geo properties - the xmax and ymax don't do anything
          # location = 'bottomright', anchor = c(x = 149, y = -18.1),                                # Position of bar
           family = "Arial", st.size = 3, st.dist = 0.3,                                            # Text properties
           border.size = 0.3, height = 0.2, st.bottom = TRUE) +                                     # Bar properties
  cowplot::draw_image('data/North_Pointer.jpg', x = 152, y = -11.5, scale = 1.4, hjust = 0.5) +
  geom_rect(aes( xmin = 144.9, xmax = 145.3, ymin = -21.45, ymax = -21), linewidth = 0.01, color = "red", fill = NA) +
    guides(fill = "none", shape = "none") 
lob_map_midrange
```






```{r}
ggsave(filename = "output/lob_map_midrange.pdf", plot = lob_map_midrange, height = 140, width = 140, device = cairo_pdf, units = "mm", dpi = 300)
```




















