---
title: "Diversity modelling - depth patterns at Lihou, Bougainville and Osprey - depth as 3 level categorical"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```

```{r}
source('code/packages.R')
source('code/helper_functions.R')
```


# We'll start with the aysymptotic species richness estimates from iNEXT.
```{r}
#load(file = "data/Rdata/lob_sprich_asymp.Rdata")
```

# Some EDA checks
```{r}
glimpse(lob_sprich_asymp)
lob_sprich_asymp %>% 
  ggplot() +
  geom_histogram(aes(x = SpecRich))

```


# Set model formula
Response variable is SpecRich
```{r}
div_depth_form <- bf(SpecRich ~ Depth + (1|Site),  family = Gamma(link = "log"))
```

# What priors do we need?
```{r}
get_prior(div_depth_form, data = lob_species_rich_asymp)
```

# What is the range of our data?
```{r}
lob_sprich_asymp |> 
  mutate(log_SR = log(SpecRich)) |> 
  group_by(Depth) |> 
  summarise(median = median(SpecRich), MAD = mad(SpecRich), 
            log_median = median(log_SR), log_MAD = mad(log_SR))
```


# Specify priors
```{r}
div_depth_cat_priors <- 
  prior(normal(3, 0.8), class = 'Intercept') +      
  prior(normal(0, 0.8), class = 'b') +
  prior(cauchy(0, 2), class = 'shape')
```

# Need to add a fraction to model with gamma
```{r}
lob_sprich_asymp <- lob_sprich_asymp %>% 
  mutate(SpecRich = SpecRich + 0.1)

```


# Fit model
```{r}
div_depth_cat_prioronlymod<- brm(formula = div_depth_form, 
            data = lob_species_rich_asymp,
            prior = div_depth_cat_priors,
            sample_prior = 'only',             
            chains = 3,
            cores = 3,             
            iter = 5000,
            warmup = 1000,  
            seed = 1234,
            thin = 4,
            silent = 0,            
            refresh = 0, 
            backend = "cmdstanr")
```



## Check
```{r}
div_depth_cat_prioronlymod |> conditional_effects() |> plot(points = TRUE, ask = FALSE)
```



# Priors look good - include data
```{r}
div_depth_cat_mod <- div_depth_cat_prioronlymod |> update(sample_prior = 'yes', control = list(adapt_delta = 0.99, max_treedepth = 20),
                                           cores = 3, seed = 1234, refresh = 0)
```


# CHeck prior - posterior influence
```{r}
div_depth_cat_mod |> SUYR_prior_and_posterior()
```


Looking good


# Chain diagnostics

## Trace plots
```{r traceplot}
div_depth_cat_mod$fit |>  stan_trace()
```
## Autocorrelation
```{r}
div_depth_cat_mod$fit |> stan_ac() 
```
## rhat
```{r}
div_depth_cat_mod$fit |> stan_rhat() 
```
## Effectiveness of sampling
```{r}
div_depth_cat_mod$fit |> stan_ess()
```


# Residuals
```{r}
resids <- make_brms_dharma_res(div_depth_cat_mod, integerResponse = FALSE)
testUniformity(resids)
plotResiduals(resids, form = factor(rep(1, nrow(lob_species_rich_asymp))))
plotResiduals(resids, quantreg = FALSE)
testDispersion(resids) # Dispersion plot (dispersion = variance/mean) - we assume mean and variance are the same, so we need to check
```


# Residuals look ok

## PP checks
```{r}
pp_check(div_depth_cat_mod, ndraws = 250)
```

# Logged
```{r}
predicted <- posterior_predict(div_depth_cat_mod)
bayesplot::ppc_dens_overlay(y = log1p(lob_species_rich_asymp$SpecRich), 
                            yrep = log1p(predicted[1:500,]))
```



```{r}
div_depth_cat_mod |> conditional_effects() |> plot(points = FALSE, ask = FALSE)

```


# Check R2
```{r}
bayes_R2(div_depth_cat_mod)
```
















# Final model
```{r}
div_depth_cat_mod_final <- div_depth_cat_mod
#save(div_depth_mod_final, "data/Rdata/div_depth_model_final.Rdata")
```
