---
title: "Diversity modelling - depth on LOB - results"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
    css: kable.css
editor_options: 
  chunk_output_type: inline  
---


```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```


# Packages and other functions
```{r}
source('code/packages.R')
source('code/extras.R')
source('code/helper_functions.R')
```


# Load model and data
```{r}
load(file = "data/Rdata/lob_species_rich_asymp.Rdata")
load(file = "data/Rdata/mod_data/div_depth_model_final.Rdata")
```

# Model summary
```{r}
div_depth_mod_final
```


# Model investigation
```{r}
div_depth_mod_final |> conditional_effects() |> plot(points = TRUE, ask = FALSE)
div_depth_mod_final |> conditional_effects() |> plot(points = FALSE, ask = FALSE)
```


## Summary table
```{r}
div_depth_mod_final |> 
  gather_draws(`^b_.*`, regex=TRUE) |> 
  #filter(.variable != 'b_Intercept') |> 
  mutate(Value = exp(.value), .keep = "unused")  |> 
  group_by(.variable) |> 
  summarise(median_hdci(Value),
            Pl = mean(Value < 1),
            Pg = mean(Value > 1))
```

## Alternative
```{r}
div_depth_mod_final |> 
  as_draws_df() |>    # Extract draws
  #select(1:4)  |> 
  exp() |>            # Back transform
  summarise_draws(median,
                  HDInterval::hdi,
                  Pl = ~mean(.x <1),                       # Exceedence prob - on exp scale it's sig if doesn't cross 1
                  Pg = ~mean(.x >1)) |> 
  as_tibble()  |> 
  mutate(across(where(is.numeric), round, 4))
```

# Summary

Strong evidence of a negative effect of depth on species richness

0.7958 * as many predators per 1SD increase in depth
-> Median = 1-0.7958 ~20% decline in predator spp richness per 1SD increase in depth
-> Max = 1-0.72 = 28%
-> Min = 1-0.87 = 13%

### Interpretation:

- Strong evidence for a negative effect of depth:
0.658 * as many predators per 1SD increase in depth (not in actual meters units) 
-> Median = 1-0.658 equivalent to a 34.2% decline per 1SD
-> Max = 1-0.578 = 42.2%
-> Min = 1-0.747 = 25.3%



## emmeans and contrasts
Will compare abundance at max and min values of both predictors

```{r}
lob_species_rich_asymp|> 
  summarise(min_Depth = min(Depth_m),
            max_Depth = max(Depth_m))
```

```{r}
lob_species_rich_asymp|> 
  summarise(mean_Depth = mean(Depth_m),
            sd_Depth = sd(Depth_m))
```


# Make dataframe with data points:
```{r}
depth_newdata <- data.frame(Depth_m = c(4, 99))
```


## emmeans for each extreme
```{r}
div_depth_mod_final |> 
  emmeans(~ Depth_m, at = depth_newdata, type = "response") 
```

5.7 species per transect in the shallows vs 2.43 at depth - half as many. Check contrast ratio


# Get contrast ratio ±HDI
```{r}
div_depth_mod_final |> 
  emmeans(~ Depth_m, at = depth_newdata) %>% 
  pairs(reverse = FALSE) |> 
  gather_emmeans_draws() |> 
  summarise(median_hdci(exp(.value), .width = c(0.8, 0.95)), 
            Pl = mean(.value < 0),
            Pg = mean(.value > 0))

```
2.4 x as many species in shallows as at depth HDCI [1.6 -	3.3]

# Contrasts
```{r}
div_depth_mod_final |> 
  emmeans(~ Depth_m, at = depth_newdata, type = "link")  |> 
  regrid() |> 
  pairs(reverse = FALSE)
```




## Figures

## Effect of depth
```{r}
depth_grid <- with(lob_species_rich_asymp, list(Depth_m = seq(from = min(Depth_m), to = max(Depth_m), len = 100)))
```

# Extract posterior draws to reconstruct trend
```{r fig.height=70/26.5, fig.width=100/25.4}
depth_div_trend <- 
div_depth_mod_final |> 
  emmeans(~ Depth_m, epred = TRUE,
           at = depth_grid) |> 
  gather_emmeans_draws() |> 
  ggplot(aes(x = Depth_m, y = .value)) +
 # geom_point(data = lob_species_rich_asymp, aes(x = Depth_m, y = SpecRich), fill = "#0066cc", colour = "midnightblue", pch = 21, alpha = 0.3, position = position_jitter(width = 0.5, height = 0.5))+
  #stat_lineribbon(colour = "black", fill = "#0066cc", size = 0.5, point_interval = "median_hdci", .width = 0.95, alpha = 0.5) +
  #stat_lineribbon(colour = "black", fill = "#0066cc", size = 0.5, point_interval = "median_hdci", .width = 0.8, alpha = 1) +
  geom_line(aes(group = .draw), colour = "#0066cc", alpha = 0.025) +
  stat_lineribbon(colour = "midnightblue", size = 0.4, point_interval = "median_hdci", fill = "#0066cc", .width = 0.001, alpha = 1) +
  scale_y_continuous(name = expression("Asymp. sp. rich. (.150m"^{2}~")"), limits = c(0,10), expand = c(0.08,0), breaks = c(0,2,4,6,8,10)) +
  scale_x_continuous("Depth (m below sea surface)") + 
  theme_bjc() +
  theme(axis.title.x = element_text(colour = "black", size = 10),
        axis.title.y = element_text(colour = "black", size = 10),
        axis.text = element_text(colour = "black", size = 9),
        legend.position = "none")

depth_div_trend
```


# Probability distribution of effect size
```{r fig.width = 70/25.4, fig.height=50/25.4}
depth_div_effect <- 
div_depth_mod_final |> 
  gather_draws(`^b_.*`, regex=TRUE)  |>    # Need to use gather_draws here if you want to use the regex
  filter(.variable == 'b_scaleDepth_m') |>
  ungroup() |> 
  ggplot(aes(x = 1-exp(.value))) + 
  stat_halfeye(fill = "#0066cc", .width = c(0.8, 0.95), point_interval = "median_hdi") +
  #t = -0.23, lty = 11) +
  geom_vline(xintercept = 0, lty = 11) +
  #annotate(geom = "text", label = "P=1", x = -0.4, y = 0.8, colour = "black", size = 3.5, family = "Arial", hjust = 0) +
  scale_y_discrete(expand = c(0,0.08)) +
  scale_x_continuous("Scaled effect distribution") +
  theme_bjc() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(colour = "black", size = 10),
        axis.text.y = element_blank(),
        axis.text.x = element_text(colour = "black", size = 9),
        legend.position = "none")

depth_div_effect
```



# Combine
```{r fig.width = 152/25.4, fig.height=60/25.4}
depth_div_fig <- 
  cowplot::plot_grid(depth_div_trend, depth_div_effect, 
                     nrow = 1, labels = c("a", "b"), rel_widths = c(1.2, 0.6), label_size = 10, align = "h") +
  theme(plot.background = element_rect(color = "black"))
depth_div_fig
```


# Save
```{r}
ggsave(filename = "output/depth_div_fig.pdf", plot = depth_div_fig, width = 152, height = 60, device = cairo_pdf, units = "mm", dpi = 300)
```




# Contrast figure: shallow to deep

# Compare across actual range of depths surveyed 4 - 99m
```{r}
depth_newdata2 <- data.frame(Depth_m = c(4, 99))
```


# Calculate contrast
```{r fig.width = 40/25.4, fig.height=50/25.4}
div_depth_contrast <- div_depth_mod_final |> 
  emmeans(~ Depth_m, at = depth_newdata2, type = "link") |> 
  pairs(reverse = FALSE) |> 
  gather_emmeans_draws() |> 
  summarise(median_hdci(exp(.value), .width = c(0.8, 0.95)),
            Pl = mean(.value < 0),
            Pg = mean(.value > 0)) |> 
  mutate(Contrast = "Diversity")
```

# Plot
```{r fig.width = 60/25.4, fig.height=35/25.4}
depth_div_full_contrast <- 
ggplot() +
  geom_pointrange(aes(y = y, ymin = ymin, ymax = ymax, x = contrast), data = div_depth_contrast |> filter(.width == 0.8), linewidth = 1.7) +
  geom_pointrange(aes(y = y, ymin = ymin, ymax = ymax, x = contrast), data = div_depth_contrast |> filter(.width == 0.95), pch = 21, size = 1.2, fill = "#0066cc", linewidth = 0.7) +
  geom_hline(yintercept = 1, lty = 11) +
  theme_bjc() +
  scale_y_continuous("Contrast ratio", position = "right", breaks = c(1,2,3,4)) +
  scale_x_discrete("Depth contrast 4-99m") +
   theme(axis.title = element_text(colour = "black", size = 10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(colour = "black", size = 9),
        legend.position = "none") +
    theme(plot.background = element_rect(color = "black", fill = NA, linewidth = 0.3))

depth_div_full_contrast
```

# Save
```{r}
ggsave(filename = "output/depth_div_full_contrast.pdf", plot = depth_div_full_contrast, width = 50, height = 50, device = cairo_pdf, units = "mm", dpi = 300)
```



# Make RHS of plot
```{r fig.width = 60/25.4, fig.height=70/25.4}
depth_div_fig_rhs <- 
  cowplot::plot_grid(depth_div_effect, depth_div_full_contrast, 
                     nrow = 2, labels = c("b", "c"), rel_heights = c(1, 1), label_size = 10) +
  theme(panel.border = element_rect(color = "black", fill = NA, linewidth = 0.3))
depth_div_fig_rhs
```

# Make full figure

```{r fig.width = 160/25.4, fig.height=80/25.4}
depth_div_fig_full <- 
  cowplot::plot_grid(depth_div_trend, depth_div_fig_rhs, 
                     nrow = 1, labels = c("a"), rel_widths = c(1, 0.6), label_size = 10) +
  theme(panel.border = element_rect(color = "black", fill = "transparent", linewidth = 0.7))
   # plot.background = element_rect(color = "black"))

# Check it looks on par with the abun fig
depth_abun_fig_full
depth_div_fig_full
```

# Save
```{r}
ggsave(filename = "output/depth_div_fig_full.pdf", plot = depth_div_fig_full, width = 160, height = 80, device = cairo_pdf, units = "mm", dpi = 300)
ggsave(filename = "output/depth_div_fig_full.jpg", plot = depth_div_fig_full, width = 160, height = 80, device = "jpeg", units = "mm", dpi = 300)
```


## EXTRAS ##


# If you want to plot abundance and diversity contrasts at the same time
```{r fig.width = 70/25.4, fig.height=50/25.4}
depth_contrasts <- div_depth_contrast |> 
  bind_rows(abun_depth_contrast)

depth_divabun_contrasts <- 
ggplot() +
  geom_pointrange(aes(y = y, ymin = ymin, ymax = ymax, x = Contrast), data = depth_contrasts |> filter(.width == 0.8), linewidth = 1.7) +
  geom_pointrange(aes(y = y, ymin = ymin, ymax = ymax, x = Contrast), data = depth_contrasts |> filter(.width == 0.95), pch = 21, size = 1.2, fill = "#0066cc", linewidth = 0.7) +
  geom_hline(yintercept = 1, lty = 11) +
  theme_bjc() +
  scale_y_continuous("Ratio", limits = c(1, 5.5)) +
  scale_x_discrete("Depth contrast (5 - 95m)") +
   theme(axis.title = element_text(colour = "black", size = 10),
        axis.text.x = element_text(colour = "black", size = 9),
        axis.text.y = element_text(colour = "black", size = 9),
        legend.position = "none")

depth_divabun_contrasts
```


# Save
```{r}
ggsave(filename = "output/depth_divabun_contrasts.pdf", plot = depth_divabun_contrasts, width = 70, height = 50, device = cairo_pdf, units = "mm", dpi = 300)
```






