---
title: "Predator diversity"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	include = FALSE,
	message = FALSE,
	warning = FALSE)
```

# Housekeeping and packages
```{r include=FALSE}
#rm(list=ls()) # Clear out environment if necessary
getwd()
source("code/packages.R")
```

# Import predator data
```{r message=FALSE, warning=FALSE, include=FALSE}
source("code/wrangling/diw_rawdata.R")
source("code/diversity/diw_divdata.R")
```


# Introduction
In this document we prepare the data in order to analyse differences in diversity between Lihou, Osprey and Bougainville and across depth gradients.
As with the wider analysis, we don't have enough observations at transect level to extract any meaningful estimate of diversity so we'll need to pool species observations by site * depth (each reef will have multiple site * depth combinations). After that we'll rarefy the data to obtain asymptotic species richness estimates.


# First extract reef, site, depth info etc.
```{r}
reef_info <- 
lob_preds |> 
  select(Reef_1, Site,
         Depth, Site_Depth) |> 
  distinct()
```

So we have 48 combinations of site and depth across our 3 reefs.


# Next, create a summary of species observations, by site - depth:
```{r}
site_depth_specsum <- 
  lob_preds %>% 
  group_by(Site_Depth, Binomial) %>% # Group including family for use in taxa_nos analysis later
  dplyr::summarise(Count=sum(Number)) %>% 
  ungroup()
site_depth_specsum

site_depth_specsum %>% 
  select(Binomial) %>% 
  unique() %>% 
  filter(!is.na(Binomial))
```


# And then convert into observation matrix

```{r}
site_depth_matrix <- 
  site_depth_specsum |> 
  pivot_wider(names_from = Binomial, values_from=Count, values_fill = 0) |> 
  dplyr::select(!`NA`) |>  # Get rid of column where the sp name ('Binomial') had been NA
  ungroup() |> 
  mutate(Total = rowSums(across(where(is.numeric)))) |> 
  filter(Total > 2)# |>                  # We are removing the 0 or 1 observation transects but will put them back in later for modelling
  select(!Total)
head(site_depth_matrix)
```

Having removed the low-observation rows, we are left with 39 site depth combos. So we have left 9 out:


# The combinations we left out were:
```{r}
site_depth_specsum |> 
  pivot_wider(names_from = Binomial, values_from=Count, values_fill = 0) |> 
  dplyr::select(!`NA`) |>  # Get rid of column where the sp name ('Binomial') had been NA
  ungroup() |> 
  mutate(Total = rowSums(across(where(is.numeric)))) |> 
  filter(Total < 3) |>                  # We are removing the 0 or 1 observation transects but will put them back in later for modelling
  select(Site_Depth, Total)
```

The totals presented here will need to be added back in to the data at the end. Note that the '2's could just be one spp:
```{r}
non.rarefied <- 
  site_depth_specsum |> 
  pivot_wider(names_from = Binomial, values_from=Count, values_fill = 0) |> 
  dplyr::select(!`NA`) |>  # Get rid of column where the sp name ('Binomial') had been NA
  ungroup() |> 
  mutate(Total = rowSums(across(where(is.numeric)))) |> 
  filter(Total < 3) |>                  # We are removing the 0 or 1 observation transects but will put them back in later for modelling
  select(!Total) |>                     # Don't want to add back in the total to the final sp count
  mutate(mutate(across(is.numeric, ~  ifelse(. > 1, 1, .)))) |> 
  mutate(SpecRich = rowSums(across(where(is.numeric)))) |>  # We'll use the same name for the iNEXT output
  select(Site_Depth, SpecRich) %>% 
  left_join(reef_info)
```




# iNEXT requires rownames as species so we'll have to flip this matrix
```{r}
inext_matrix <- 
  site_depth_matrix |> 
  column_to_rownames("Site_Depth") |> 
  t() #|> 
```


# Run in iNEXT
```{r}
site_depth_rarefied <- 
  inext_matrix |> 
  iNEXT(q = c(0,1),  datatype = "abundance", endpoint = 1000, nboot = 100)
```


# Explore output
```{r}
# Explore
site_depth_rarefied$DataInfo # basic data information.

site_depth_rarefied$AsyEst # asymptotic diversity estimates.
site_depth_rarefied$iNextEst$size_based 
site_depth_rarefied$iNextEst$coverage_based 

```


# Extract just the asymptotic species richness and combine with the other reef info
```{r}
lob_sprich_asymp <-
  reef_info %>% full_join(
  site_depth_rarefied$AsyEst  |> 
               rename(Site_Depth = Assemblage,
                      SpecRich = Estimator) |>
               filter(Diversity == "Species richness")) %>% 
  filter(!is.na(Diversity)) %>% 
  bind_rows(non.rarefied) # Add back in our non-rarefied 0/1 observations
```



# NEED TO BOOTSTRAP HERE #

```{r}
nest()

install.packages("rsample")



```




# Quick EDA

### By reef
```{r fig.width = 150/25.4, fig.height = 90/25.4}
lob_sprich_asymp %>% 
  group_by(Reef_1) %>% 
  summarise_at(vars(SpecRich), list(mean = mean, SD = sd, SE = std.error)) %>% 
  ggplot() + 
  geom_pointrange(aes(x = Reef_1, y = mean, ymin = mean-SE, ymax = mean+SE)) +
  geom_point(aes(x = Reef_1, y = mean)) +
  theme_minimal() +
  ylab("Diversity (no of spp per site ± SE)") +
  xlab("Reef")
```


### Depth
```{r fig.width = 150/25.4, fig.height = 90/25.4}
lob_sprich_asymp %>% 
  group_by(Depth) %>% 
  summarise_at(vars(SpecRich), list(mean = mean, SD = sd, SE = std.error)) %>% 
  ggplot() + 
  geom_pointrange(aes(x = Depth, y = mean, ymin = mean-SE, ymax = mean+SE)) +
  geom_point(aes(x = Depth, y = mean)) +
  geom_line(aes(x = Depth, y = mean, group = 1)) +
  theme_minimal() +
  ylab("Diversity (no of spp per site ± SE)") +
  xlab("Depth zone")
```


# Reef * Depth
```{r}
lob_sprich_asymp %>% 
  group_by(Reef_1, Depth) %>% 
  summarise_at(vars(SpecRich), list(mean = mean, SD = sd, SE = std.error)) %>% 
  ggplot() + 
  geom_pointrange(aes(x = Depth, y = mean, ymin = mean-SE, ymax = mean+SE)) +
  geom_point(aes(x = Depth, y = mean)) +
  geom_line(aes(x = Depth, y = mean, group = 1)) +
  theme_minimal() +
  ylab("Diversity (no of spp per site ± SE)") +
  xlab("Depth zone") +
  facet_wrap(~ Reef_1)

```















