---
title: "Abundance wrangling"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
```


# Load predator data, packages
```{r message=FALSE, warning=FALSE, include=FALSE, fig.width = 150/25.4, fig.height = 90/25.4}
# Clear out 
rm(list=ls()) 

# Load packages, extras and helper functions
sourceRmd::sourceRmd("code/0_extras/source_extras.Rmd")

# Load data
# Either
#source("code/wrangling/diw_rawdata.R")
# Or
load('data/2_Rdata/lob_preds.Rdata')
```


```{r message=FALSE, warning=FALSE, include=FALSE, fig.width = 150/25.4, fig.height = 90/25.4}
# Check data
lob_preds

# Distribution of response variable - per transect
lob_preds |> group_by(T_ID) |> 
  summarise(Transect_Abun = sum(Number)) |> 
  ggplot() +  
  geom_histogram(aes(x=Transect_Abun), binwidth = 1, color="black", fill="grey") 
  scale_x_continuous("No. of predators")

# Per observation
lob_preds  |>  ggplot() +  
  geom_histogram(aes(x=Number), binwidth = 1, color="black", fill="grey") +
  scale_x_continuous("No. of predators")
```


```{r}
lob_preds |> group_by(T_ID) |> 
  summarise(Transect_Abun = sum(Number)) |> 
  filter(Transect_Abun == 0) |> 
  tally()
```


Quite a few zeros (98 = almost 1/3 of transects) and definitely right skewed. Indicating hurdle or ZI model

## Let's check what our ranges are
```{r}
lob_preds  |> 
  summarise_at(vars(Number), list(median = median, mean = mean, SD = sd, SE = std.error,
                                  CI5 = ~quantile(., probs = 0.05), CI95 = ~quantile(., probs = 0.95), 
                                  CI1 = ~quantile(., probs = 0.01), CI99 = ~quantile(., probs = 0.99), 
                                  Q1 = ~quantile(., probs = 0.25), Q3=~quantile(., probs = 0.75)))
```

# In summary:

The average number of preds observed per transect is around 1 (median and mean) 
95% of observations are 3 or less and a lot of zeros
Fair bit of variability (SD is nearly double the mean)
1% of observations are 12 or more. We could round down these outliers to minimise their effect, however these are likely schooling species, so probably worth keeping in.

## See how many observations removing outliers would affect:
```{r}
lob_preds |>  filter(Number >12)
```
  
This is basically 8 schools of fishes observed over the study. Species look appropriate - we will leave these alone for modelling

# Next we need a tbl that presents counts of predators per transect:
```{r}
lob_predsum <- 
   lob_preds |> 
   group_by(Reef_1, Reef_2, Site,                         # Grouping variables of reef and site - only going to need Reef_1
            Depth, Depth_m, Depth_bin, Depth_bin_meso,    # Can model depth as continuous or as a binned variable  
            #Site_lat, Site_long,                         # Don't need these vars for this analysis (see BioGeo repo for more info)
            #dCT, dGBR, Region,                           # "
            #Reef_1_Area, Reef_size,                      # "
            #Isol, IsolRank, Reef_100,                    # " 
            #Aspect_descriptive, Aspect_scale,            # "
            #Temperature, Benthic_Complexity,             # "
            #Situation, Habitat, Zone,
            T_ID) |>
   summarise(No_Preds = sum(Number)) |>  #
   ungroup()
```

# Now have a tbl - 346 obs/T_ID:
```{r}
lob_predsum
```

# Check that agrees with what we calculated from the main tbl:
```{r}
lob_preds |>  select(T_ID) |> distinct()
```

Yep

# Save if required
```{r}
save(lob_predsum, file = 'data/2_Rdata/lob_predsum.Rdata')
```


# Checks

# Total # of predators observed (again, use both tbls to check)
```{r}
lob_predsum  |> summarise(Total_preds = sum(No_Preds))
lob_preds |> summarise(No_Preds = sum(Number)) 
```


## Transects per reef
```{r}
lob_predsum  |> 
  group_by(Reef_1) |> 
  summarise(Transect_number = n())
```
Slightly unbalanced between reefs, but at least in the right order - the largest has the most transects and vice versa

## How many sites?
```{r}
lob_predsum  |> 
  select(Reef_1, Site) |> 
  distinct() |> 
  group_by(Reef_1) |> 
  summarise(No_sites = n())
```
Lihou has many more sites due to how spread out it is


## Sample size - by 10m depth bin
```{r}
lob_predsum |> 
  group_by(Depth_bin) |> 
  summarise(Transect_number = n())
```
Reasonably balanced

## Sample size - by depth zone
```{r}
lob_predsum |> 
  group_by(Depth_bin_meso) |> 
  summarise(Transect_number = n())
```
Reasonably balanced



## Predator abundance by reef

### We know from the parallel analysis on reef isolation/biogeography that smaller reefs have more predators so this isn't that surprising:
```{r fig.width = 150/25.4, fig.height = 90/25.4}
lob_predsum  |> 
  group_by(Reef_1) |> 
  summarise_at(vars(No_Preds), list(median = median, mean = mean, SD = sd, SE = std.error)) |> 
  ggplot() +
  geom_pointrange(aes(x = Reef_1, y = mean, ymin = mean-SE, ymax = mean+SE)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

# Data exploration

## Overview of depth effect?

### As continuous
```{r fig.width = 150/25.4, fig.height = 90/25.4}
lob_predsum |> 
  ggplot() + 
  geom_point(aes(x = Depth_m, y = No_Preds)) +
  geom_smooth(aes(x = Depth_m, y = No_Preds), method = "glm") +
  theme_minimal() +
  ylab("No of predators per transect") +
  xlab("Depth (m)")
```


A lot of noise but looks like decreasing abundance with increasing depth


### 10m depth bins
```{r fig.width = 150/25.4, fig.height = 90/25.4}
lob_predsum  |> 
  group_by(Depth_bin) |> 
  summarise_at(vars(No_Preds), list(mean = mean, SD = sd, SE = std.error)) |> 
  ggplot() + 
  geom_pointrange(aes(x = Depth_bin, y = mean, ymin = mean-SE, ymax = mean+SE)) +
  geom_point(aes(x = Depth_bin, y = mean)) +
  geom_line(aes(x = Depth_bin, y = mean, group = 1)) +
  theme_minimal() +
  ylab("Abundance (no of predators ± SE)") +
  xlab("Depth (10m bins)")
```


### Photic zone bins
```{r fig.width = 150/25.4, fig.height = 90/25.4}
lob_predsum |> 
  group_by(Depth) |> 
  summarise_at(vars(No_Preds), list(mean = mean, SD = sd, SE = std.error)) |> 
  ggplot() + 
  geom_pointrange(aes(x = Depth, y = mean, ymin = mean-SE, ymax = mean+SE)) +
  geom_point(aes(x = Depth, y = mean)) +
  geom_line(aes(x = Depth, y = mean, group = 1)) +
  theme_minimal() +
  ylab("Abundance (no of predators ± SE)") +
  xlab("Depth zone")
```


