---
title: "Asymtotic species richness wrangling"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
```


# Load predator data, packages
```{r message=FALSE, warning=FALSE, include=FALSE, fig.width = 150/25.4, fig.height = 90/25.4}
# Clear out 
rm(list=ls()) 

# Load packages, extras and helper functions
sourceRmd::sourceRmd("code/0_extras/source_extras.Rmd")

# Load data
# Either
#source("code/wrangling/diw_rawdata.R")
# Or
load('data/2_Rdata/lob_preds.Rdata')
load("data/2_Rdata/lob_predsum.Rdata")
load("data/2_Rdata/lob_metadata.Rdata")
```


# Introduction
We have 346 transects over 17 sites, with pretty even depth coverage:
```{r}
lob_metadata |> select(T_ID) |> distinct()
lob_metadata |> select(Site) |> distinct()
lob_metadata |> select(Depth) |> summary()
```

However, there are quite a few transects with few or 0 observations
```{r}
nopreds <- lob_preds |> 
  select(Reef_1, T_ID, Binomial) |> 
  distinct()  |> 
  filter(is.na(Binomial)) |> 
  select(Reef_1, T_ID) |> 
  mutate(n = 0)
```
98 with no predators

We are going to rarefy these data to get asymptotic species richness estimates, but will have to manage these observations separately.

# Run rarefaction script
```{r}
source('code/3_sprich/transect_rarefaction.R')
```


## The asymptotic species richness is in a sub-set of the iNEXT object
```{r}
transect_rarefied$AsyEst
```

This has each of the following per transect:
- Species richness 
- Shannon diversity
- Simpson diversity 
- plus SE and CI for each of the above

# We can extract just species richness from this and combine with our predictor variables (env_vars)
```{r}
species_rich_asymp <- 
  env_vars |> 
  left_join(transect_rarefied$AsyEst |> 
  rename(T_ID = Assemblage)) |>
  mutate_if(is.character, as.factor) |> 
  filter(Diversity == "Species richness") |> 
  rename(SpecRich = Estimator)
```


## Check
```{r}
species_rich_asymp
```
Of course we only have 248 transects in here as there are 98 with no observations that were excluded from rarefaction

## Add back in 0 obs transects by full_join with env_vars
```{r}
species_rich_asymp <- 
  env_vars |> 
  left_join(transect_rarefied$AsyEst |> 
  rename(T_ID = Assemblage)) |>
  mutate_if(is.character, as.factor) |> 
  filter(Diversity == "Species richness") |> 
  rename(SpecRich = Estimator)  |> 
  full_join(env_vars) |> 
  mutate(across(where(is.double), ~replace_na(., 0)))
```


# And save file for modelling
```{r}
save(species_rich_asymp, file = "data/Rdata/lob_species_rich_asymp.Rdata")
```
